### 2024-11-04 artefacts/paper.tex
add paper tex, add max gradient perturb in readme

\documentclass{article}
\usepackage{graphicx} % Required for inserting images

\title{MultifidelityTrajectoryOptimization}
\author{goroda }
\date{October 2024}

\begin{document}

\maketitle

\section{Introduction}

\end{document}

### 2024-11-04 stoch_traj.m

@@ -199,34 +199,6 @@ for idx = 1:length(TsList)
 end
 
 %% *** correlation check for MLMC, ONLY FOR 2 LEVEL
-%% correlation at different points in a numerical optimizer in HF
-Uopt_hf = data.lqrsol{1}.Uopt;
-% u0_num = zeros(size(Uopt_hf));
-u0_num = repelem(data.lqrsol{2}.Uopt, 10, 1); % warm start
-global num_opt_iters num_opt_data;
-num_opt_iters = 0;
-num_opt_data = zeros(size(Uopt_hf,1), 100);
-fun = @(u) LQRCostwithGrad(data.lqrsol{1}.x0_ext, u, data.lqrsol{1}.Q_ext, data.lqrsol{1}.S, data.lqrsol{1}.M, data.lqrsol{1}.Qbar, data.lqrsol{1}.Rbar);
-options = optimoptions('fminunc', 'Display', 'iter', 'Algorithm', 'quasi-newton', 'SpecifyObjectiveGradient', true, 'OutputFcn', @outfun);
-Uopt_num = fminunc(fun, u0_num, options);
-uopt_diff = sqrt(sum((Uopt_hf - Uopt_num).^2));
-
-U_hf = num_opt_data(:, 1:num_opt_iters);
-% U_lf = repelem(U_hf(1:10:end, :), 10, 1);
-U_lf = downsample_avg(U_hf, 10); % get LF by avg every 10 values of HF
-vis_sols(U_hf, U_lf, data.lqrsol{1}.times, "Solutions along optimizer path", 1:num_opt_iters, "Iteration");
-% calculate correlation along the optimizer path
-[cost_hf, cost_lf] = calc_costs_multifid(data.lqrsol{1}, U_hf, U_lf);
-corr = calc_corr_multifid(x0_rv, u0, ref, data.lqrsol{1}, U_hf, U_lf);
-% TODO: fix title of this graph
-plot_multifid_costs(1:num_opt_iters, cost_hf, cost_lf, corr, "optimizer path", "Iteration");
-
-% Get correlation for all iterations
-corr = calc_corr_multifid_2d_iters(x0_rv, u0, ref, data.lqrsol{1}, U_hf, U_lf);
-
-%% plot correlation
-plot_corr_2d(corr, "Correlation between HF/LF costs of solutions at different iterations in HF", "hf_iters_corr");
-
 %% *** correlation with perturbations
[…]

### 2024-11-04 stoch_traj.m
add separate sims in cost function calcs, add basic mlmc estimator

@@ -4,10 +4,10 @@ clc; clear; close all force;
 % system params
 m = 5; % mass
 k = 2; % spring coefficient
-b = 0.5; % damping coefficient
+c = 0.5; % damping coefficient
 
 A_msd = [0 1;
-  -k / m -b / m];
+  -k / m -c / m];
 B_msd = [0;
   1 / m];
 % C_msd = [1 0]; % only observe position
@@ -212,7 +212,7 @@ vis_sols(Uopt_hf, Uopt_lf, data.lqrsol{1}.times, "HF sim", perturb_range, "Pertu
 
 % calculate costs and correlation for both high and low fidelity, for each perturbation
 [cost_hf, cost_lf] = calc_costs_multifid(data.lqrsol{1}, Uopt_hf, Uopt_lf);
-corr = calc_corr_multifid(x0_rv, u0, ref, data.lqrsol{1}, Uopt_hf, Uopt_lf);
+corr = calc_corr_multifid(x0_rv, u0, ref, data.lqrsol{1}, data.lqrsol{1}, Uopt_hf, Uopt_lf);
 %% plot costs
 plot_multifid_costs(perturb_range, cost_hf, cost_lf, corr, "HF sim", "Perturbation");
 
@@ -228,7 +228,7 @@ vis_sols(Uopt_hf, Uopt_lf, data.lqrsol{2}.times, "LF sim", perturb_range, "Pertu
 
 % calculate costs and correlation for both high and low fidelity, for each perturbation
 [cost_hf, cost_lf] = calc_costs_multifid(data.lqrsol{2}, Uopt_hf, Uopt_lf);
-corr = calc_corr_multifid(x0_rv, u0, ref, data.lqrsol{2}, Uopt_hf, Uopt_lf);
+corr = calc_corr_multifid(x0_rv, u0, ref, data.lqrsol{2}, data.lqrsol{2}, Uopt_hf, Uopt_lf);
 
 %% plot costs
 plot_multifid_costs(perturb_range, cost_hf, cost_lf, corr, "LF sim", "Perturbation");
@@ -250,19 +250,44 @@ U_hf = num_opt_data(:, 1:num_opt_iters);
 U_lf = downsample_avg(U_hf, 10); % get LF by avg every 10 values of HF
[…]

### 2024-11-05 stoch_traj.m
fix plots and fix mlmc estimator

@@ -1,5 +1,5 @@
 clc; clear; close all force;
-%% state space model
+%% A. state space model
 
 % system params
 m = 5; % mass
@@ -48,7 +48,7 @@ perturb_dir_samples = 1000;
 perturb_dir_mag = 0.1;
 perturb_range = -1:0.25:6;
 
-%% simulate open loop dynamics
+%% B. open loop dynamics
 [data.t, x_ode] = ode45(@(t, x) msd(t, x, 0, A_msd, B_msd), 0:Tfid:Tsim, x0);
 data.x_ol = x_ode';
 
@@ -65,7 +65,7 @@ legend show; legend boxoff;
 grid on; grid minor;
 saveas(fig, 'figs/ol_det.svg');
 
-%% stochastic with random initial state
+%% C. stochastic dynamics with random initial state
 
 % set up problem with stochastic initial state
 x0_rv = mvnrnd(x0_mean, x0_cov, rv_samples)';
@@ -74,14 +74,14 @@ x0_rv_cov = cov(x0_rv');
 x0_rv_mean_ext = [x0_rv_mean; u0; ref];
 x0_rv_cov_ext = blkdiag(x0_rv_cov, zeros(3,3));
 
-% simulate open loop dynamics
+%% C.1 open loop dynamics
 data.x_ol_stoch = zeros(nx, Tsim/Tfid + 1, rv_samples);
 for i = 1:rv_samples
   [~, x_ode] = ode45(@(t, x) msd(t, x, 0, A_msd, B_msd), 0:Tfid:Tsim, x0_rv(:, i));
   data.x_ol_stoch(:, :, i) = x_ode';
 end
 
-% plot open loop dynamics
+%% C.2 plot open loop dynamics
 fig = figure;
 hold on;
 
@@ -99,6 +99,8 @@ legend show; legend boxoff;
 grid on; grid minor;
 saveas(fig, 'figs/ol_stoch_init.svg');
 
+%% D closed loop dynamics with LQR
+
 data.lqrsol = cell(length(TsList), 1);
 
 for idx = 1:length(TsList)
@@ -106,7 +108,7 @@ for idx = 1:length(TsList)
   fprintf("\n*** Ts: %.2f ***\n", Ts);
   assert(mod(Ts, Tfid) == 0, "Ts=" + Ts + " is not a multiple of Tfid=" + Tfid);
   
-  %% deterministic LQR
+  %% D.1 deterministic LQR
   Tmult = Ts / Tfid;
   times = 0:Ts:Tsim;
[…]

### 2024-11-11 stoch_traj.m
Intermediate step before doing things correctly

@@ -62,7 +62,7 @@ ylim(plot_ylim_l);
 title("Open loop dynamics");
 xlabel('Time [s]');
 legend show; legend boxoff;
-grid on; grid minor;
+grid on;
 saveas(fig, 'figs/ol_det.svg');
 
 %% C. stochastic dynamics with random initial state
@@ -96,7 +96,7 @@ ylim(plot_ylim_l);
 title("Open loop dynamics with stochastic initial state (" + rv_samples + " samples)");
 xlabel('Time [s]');
 legend show; legend boxoff;
-grid on; grid minor;
+grid on;
 saveas(fig, 'figs/ol_stoch_init.svg');
 
 %% D closed loop dynamics with LQR
@@ -139,7 +139,7 @@ for idx = 1:length(TsList)
   title("(Ts:"+Ts+") Closed loop dynamics");
   xlabel('Time [s]');
   legend show; legend boxoff;
-  grid on; grid minor;
+  grid on;
   saveas(fig, "figs/"+Ts+"_cl_det.svg");
   
   %% D.2 stochastic LQR using robust SAA
@@ -176,7 +176,7 @@ for idx = 1:length(TsList)
   title("(Ts:"+Ts+") Closed loop dynamics with stochastic initial state (" + rv_samples + " samples)");
   xlabel('Time [s]');
   legend show; legend boxoff;
-  grid on; grid minor;
+  grid on;
   saveas(fig, "figs/"+Ts+"_cl_stoch_init.svg");
   
   
@@ -196,7 +196,7 @@ for idx = 1:length(TsList)
   title("(Ts:"+Ts+") Cost distribution of stochastic LQR ("+rv_samples+" samples)");
   xlabel('Cost');
   legend show; legend boxoff;
-  grid on; grid minor;
+  grid on;
   saveas(fig, "figs/"+Ts+"_cost_dist.svg");
 end
 
@@ -207,6 +207,7 @@ end
 perturbation = get_perturb_max_grad(data.lqrsol{1}, perturb_dir_samples, perturb_dir_mag, perturb_range);
 perturb_lf = downsample_avg(perturbation, 10); % downsample to low fid while keeping the same no of elements
 
+%% E.1.1.1 in HF obj
[…]

### 2024-11-11 artefacts/paper.tex
fix correlations and graphs

@@
 \documentclass{article}
+\usepackage[a4paper, total={7in, 8in}]{geometry}
 \usepackage{graphicx} % Required for inserting images
+\usepackage{amsmath} % Required for some math elements
+\usepackage{hyperref} % Required for hyperlinks
 
@@
 
+\begin{abstract}
+  TODO: abstract
+\end{abstract}
+
 \section{Introduction}
+TODO: introduction
+\section{Background}
+TODO: background
+
+\section{Problem Formulation}
+\subsection{Mass-Spring-Damper System}
+For this paper, we consider a simplified trajectory optimization problem that is easy to understand and analyze.
+We try to solve a reference tracking problem for the position and velocity of a simple mass-spring-damper system.
+The mass-spring-damper system has the following parameters:
+\begin{itemize}
+  \item Mass: $m = 5$ g
+  \item Spring constant: $k = 2$ N/m
+  \item Damping coefficient: $c = 0.5$ Ns/m
+    % \item Initial position: $x_0 = 0$ m
+    % \item Initial velocity: $\dot{x}_0 = 0$ m/s
+\end{itemize}
+
+Given an external force $F$ and the acceleration of the system $\ddot{x}$,
+the dynamics of the system can be written as the equality of forces acting on the mass:
+
+\begin{equation}
+  m\ddot{x} = F - c\dot{x} - kx
+  \implies \ddot{x} = \frac{F}{m} - \frac{c}{m}\dot{x} - \frac{k}{m}x
+\end{equation}
+
+We track the position $x$ and velocity $\dot{x}$ of the mass-spring-damper system using the control input $u$, which applies the force $F$ on the mass.
+These are the states and control inputs of the system, giving use the following dynamics
+
+\begin{equation}
+  \begin{bmatrix}
+    \dot{x} \\
+    \ddot{x}
+  \end{bmatrix} =
+  \begin{bmatrix}
+    0 & 1 \\
+    -\frac{k}{m}  -\frac{c}{m}
+  \end{bmatrix}
+  \begin{bmatrix}
+    x \\
+    \dot{x}
+  \end{bmatrix} +
+  \begin{bmatrix}
+    0 \\
+    \frac{1}{m}
+  \end{bmatrix} u
+\end{equation}
+
+For a closed-loop system, we need to observe the output $y$ of the system, and in this case we observe the position $x$.
+
+\begin{equation}
+  y =
+  \begin{bmatrix}
+    1 & 0
+  \end{bmatrix}
+  \begin{bmatrix}
+    x \\
+    \dot{x}
+  \end{bmatrix}
+\end{equation}
+
+This gives us an LTI system in the form of $\dot{x} = Ax + Bu$ and $y = Cx$, where
+
+\begin{equation}
+  \begin{aligned}
+    A &=
+    \begin{bmatrix}
+      0 & 1 \\
+      -\frac{k}{m} & -\frac{c}{m}
+    \end{bmatrix} \\
+    B &=
+    \begin{bmatrix}
+      0 \\
+      \frac{1}{m}
+    \end{bmatrix} \\
+    C &=
+    \begin{bmatrix}
+      1 & 0
+    \end{bmatrix}
+  \end{aligned}
+\end{equation}
+
+\subsection{LQR Controller}
+We set up an LQR controller to track the reference position and velocity of the mass-spring-damper system.
+An LQR controller was chosen because it has a quadratic cost function that can be minimized analytically and gives us a
+good reference for the trajectory optimization problem.
+
+To set up the LQR controller, we need to define the cost function of the system, which consists of:
+
+\begin{itemize}
+  \item $u_0, u_1 ... u_{N-1}$ : control inputs at each timestep of the trajectory
+  \item $J$ : total cost to be minimized
+  \item $N$ : time horizon
+  \item $x_k$ : state vector at time step $k$
+  \item $u_k$ : control input vector at time step $k$
+  \item $\Delta u_k = u_k - u_{k-1}$ : change in control input at time step $k$
+  \item $r_k$ : reference output at time step $k$
+  \item $e_k = y_k - r_k = Cx - r_k$ : error in the output at time step $k$
+  \item $Q_e$ : error cost matrix
+  \item $R$ : control cost matrix
+\end{itemize}
+The LQR controller minimizes the cost function
+
+\begin{equation}
+  \min_{u_0, u_1 ... u_{N-1}}J = \sum_{k=0}^{N-1} (e_k^T Q_e e_k + \Delta u_k^T R \Delta u_k)
+\end{equation}
+
+\section{Control Variates}
+
+\subsection{Comparing optimums of high and low fidelity models} \label{comp_opt}
+
+To determine if the low fidelity model can be used as a control variate, we need to determine if it coreelates with the high fidelity model.
+
+System Setup:
+\begin{itemize}
+  \item High fidelity model/cost function: $J_h(u, x_0)$
+  \item Low fidelity model/cost function: $J_l(u, x_0)$
+  \item Control input in high fidelity model: $u_h$
+  \item Size of $u_h$: $n_{u_h} \times 1$
+  \item Control input in low fidelity model: $u_l$
+  \item Size of $u_l$: $n_{u_l} \times 1$
+  \item Random initial state: $x_0$
+  \item High fidelity optimum: $u_h^*$
+  \item Low fidelity optimum: $u_l^*$
+\end{itemize}
+
+$u_h$ and $u_l$ do not have the same dimensions, so to compare them we need to scale them to the same dimensions.
+
+To upsample $u_l \rightarrow u_h$, we repeat the elements of $u_l$ to match the dimensions of $u_h$.
+This can be seen as multiplying $u_l$ with a matrix $T_{lh}$ of size $n_{u_h} \times n_{u_l}$:
+
+\begin{equation}
+  T_{lh} =
+  \begin{bmatrix}
+    1 & 0 & \ldots & 0 \\
+    1 & 0 & \ldots & 0 \\
+    \vdots & \vdots & \ddots & \vdots \\
+    0 & 1 & \ldots & 0 \\
+    0 & 1 & \ldots & 0 \\
+    \vdots & \vdots & \ddots & \vdots \\
+    0 & 0 & \ldots & 1 \\
+    0 & 0 & \ldots & 1 \\
+    \vdots & \vdots & \ddots & \vdots \\
+  \end{bmatrix}
+\end{equation}
+
+where the number of repeating ones in each column is determined by the ratio $r_{hl} = n_{u_h} / n_{u_l}$.
+This gives us $u_{lh} = T_{lh} u_l$, which can be compared with $u_h$.
+
+To downsample $u_h \rightarrow u_l$, we can either restrict the dimensions of $u_h$ to match $u_l$ or average the elements of $u_h$ to match the dimensions of $u_l$.
+
+In case of restriction, we can multiply $u_h$ with a matrix $T_{hlr}$ of size $n_{u_l} \times n_{u_h}$:
+
+\begin{equation}
+  T_{hlr} =
+  \begin{bmatrix}
+    1 & 0 & \ldots & 0 & 0 \\
+    0 & \ldots & 1 & \ldots & 0 \\
+    \vdots & \ddots & \vdots & \ddots & \vdots \\
+    0 & 0 & \ldots & 0 & 1 \\
+  \end{bmatrix}
+\end{equation}
+
+Where the number of columns after which the next one is selected is determined by the ratio $r_{hl} = n_{u_l} / n_{u_h}$.
+This gives us $u_{hlr} = T_{hlr} u_h$, which can be compared with $u_l$.
+
+To downsample $u_h$ to $u_l$ by averaging, we need the ratio $r_{lh} = n_{u_h} \ n_{u_l}$.
+We can then multiply $u_h$ with a matrix $T_{hla}$ of size $n_{u_l} \times n_{u_h}$:
+
+\begin{equation}
+  T_{hla} =
+  \begin{bmatrix}
+    r_{lh} & \ldots & 0 & 0 & 0 \\
+    0 & \ldots & r_{lh} & \ldots & 0 \\
+    \vdots & \ddots & \vdots & \vdots & \vdots \\
+    0 & 0 & 0 & r_{lh} & \ldots \\
+  \end{bmatrix}
+\end{equation}
+
+Where $r_{lh}$ is repeated $r_{hl}$ times in each row.
+This gives us $u_{hla} = T_{hla} u_h$, which can be compared with $u_l$.
+
+\subsection{Correlation between high and low fidelity models}
+
+$u_h$ and $u_l$ each have $n_{u_h}$ and $n_{u_l}$ dimensions respectively, which represent the time steps of the simulation.
+After converting them to the same dimensions as done in section \ref{comp_opt},
+we can compute the cost for some realizations of the initial state $x_0$.
+This will give us a set of costs of both high and low fidelity control inputs for a certain model/cost function.
+We can then compute the correlation coefficient $\rho$ between the costs.
+To illustrate this, assume we have $N$ realizations of the initial state $x_0$, and J is the cost function.
+$u_h$ and $u_l$ are the control inputs for the high and low fidelity models respectively, where the dimensions have been
+converted to the same number as described in section \ref{comp_opt}. $\mu_h$ and $\mu_l$ are the mean costs
+of the high and low fidelity control inputs respectively, and $\sigma_h$ and $\sigma_l$ are the standard deviations of
+the costs of the high and low fidelity control inputs respectively.
+
+\begin{equation}
+  \rho = \frac{1}{N-1} \sum_{i=1}^{N} \left( \frac{J(u_h, x_0^i) - \mu_h}{\sigma_h} \right) \left( \frac{J(u_l, x_0^i) - \mu_l}{\sigma_l} \right)
+\end{equation}
+
+This correlation coefficient $\rho$ can be calculated using the MATLAB function \href{https://www.mathworks.com/help/matlab/ref/corrcoef.html#bunkaln}{\texttt{corrcoef}}.
+
+\subsection{Correlation Experiments}
+
+We conducted experiments to check the correlation in 2 scenarios:
+\begin{itemize}
+  \item Perturbation in the direction of steepest ascent
+  \item Along the path of a numerical optimizer
+\end{itemize}
+
+\subsubsection{Perturbation in the direction of steepest ascent}
+
+Since the solution of the cost function is determined analytically, we cannot use it to determine the direction of steepest ascent.
+The gradient at the solution will be 0. To get the direction of steepest ascent, the solution was perturbed slightly in random directions and the direction with the highest gradient
+was chosen as the direction of steepest ascent.
+
+\textbf{Perturbation in the high fidelity model/cost function}
+
+When perturbing in the high fidelity model/cost function, the perturbed high fidelity solution was downsampled to low fidelity
+dimensions by either restricting or averaging as described in section \ref{comp_opt}.
+
+Hence, we have the following important variables:
+
+\begin{itemize}
+  \item High fidelity model/cost function: $J_h$
+  \item Low fidelity model/cost function: $J_l$
+  \item High fidelity optimum: $u_h^*$
+  \item High fidelity perturbation: $p_h$
+  \item Perturbed high fidelity solution: $u_h = u_h^* + p_h$
+  \item Perturbed high fidelity solution downsampled using restriction: $u_{hlr} = T_{hlr} u_h$
+  \item Perturbed high fidelity solution downsampled using averaging: $u_{hla} = T_{hla} u_h$
+\end{itemize}
+
+Comparing $J_h(u_h, x_0)$ and $J_l(u_{hlr}, x_0)$
+
+Comparing $J_h(u_h, x_0)$ and $J_l(u_{hla}, x_0)$
+
+\textbf{Perturbation in the low fidelity model/cost function}
+
+When perturbing in the low fidelity model/cost function, the perturbed high fidelity solution was determined
+by converting the low fidelity perturbation to high fidelity dimensions as described in section \ref{comp_opt}.
+
+Hence, we have the following important variables:
+
+\begin{itemize}
+  \item High fidelity model/cost function: $J_h$
+  \item Low fidelity model/cost function: $J_l$
+  \item Low fidelity optimum: $u_l^*$
+  \item Low fidelity perturbation: $p_l$
+  \item Perturbed low fidelity solution: $u_l = u_l^* + p_l$
+  \item Perturbed high fidelity solution upsampled: $u_{lh} = T_{lh} u_l$
+\end{itemize}
+
+Comparing $J_h(u_{lh}, x_0)$ and $J_l(u_l, x_0)$
+
+\subsubsection{Along the path of a numerical optimizer}
+
+Since in a realistic scenario, one would be optimizing using a numerical optimizer, and desire to reach the high fidelity optimum,
+we conducted experiments to check the correlation along the path of a numerical optimizer in the high fidelity model/cost function.
+
+Here, we warm start with the low fidelity optimum, and check the correlation by downsampling the high fidelity solution using averaging.
+
+Hence, we have the following important variables:
+
+\begin{itemize}
+  \item High fidelity model/cost function: $J_h$
+  \item Low fidelity model/cost function: $J_l$
+  \item High fidelity solution: $u_h$
+  \item High fidelity solution downsampled using restriction: $u_{hlr} = T_{hlr} u_h$
+  \item High fidelity solution downsampled using averaging: $u_{hla} = T_{hla} u_h$
+\end{itemize}
+
+Comparing $J_h(u_h, x_0)$ and $J_h(u_{hlr}, x_0)$
+
+Comparing $J_h(u_h, x_0)$ and $J_l(u_{hla}, x_0)$
 

### 2024-11-11 stoch_traj.m

@@ -200,71 +200,50 @@ for idx = 1:length(TsList)
   saveas(fig, "figs/"+Ts+"_cost_dist.svg");
 end
 
-%% *** E. correlation check for MLMC, ONLY FOR 2 LEVEL TODO: extend to work for any number of levels
-%% E.1 correlation with perturbation in direction of maximum ascent
+% *** E. correlation check for MLMC, ONLY FOR 2 LEVEL TODO: extend to work for any number of levels
+% E.1 correlation with perturbation in direction of maximum ascent
 
 %% E.1.1 perturbation in HF objective fn
 perturbation = get_perturb_max_grad(data.lqrsol{1}, perturb_dir_samples, perturb_dir_mag, perturb_range);
-perturb_lf = downsample_avg(perturbation, 10); % downsample to low fid while keeping the same no of elements
-
-%% E.1.1.1 in HF obj
 Uopt_hf = data.lqrsol{1}.Uopt + perturbation; % perturb hf sol
-Uopt_lf = repelem(data.lqrsol{2}.Uopt + perturb_lf(1:10:end, :), 10, 1); % pick every 10th elem and repeat 10 times
-vis_sols(Uopt_hf, Uopt_lf, data.lqrsol{1}.times, "HF obj", perturb_range, "Perturbation");
 
-% calculate costs and correlation for HF/LF sols for each perturbation
-[cost_hf, cost_lf] = calc_costs_multifid(data.lqrsol{1}, data.lqrsol{1}, Uopt_hf, Uopt_lf);
-corr = calc_corr_multifid(x0_rv, u0, ref, data.lqrsol{1}, data.lqrsol{1}, Uopt_hf, Uopt_lf);
-%% plot costs
-title_str = ["$J_h$", "$J_h$"];
-costs_str = ["$J_h(u_h)$", "$J_h(u_{lh})$"];
-plot_multifid_costs(perturb_range, cost_hf, cost_lf, corr, title_str, costs_str, "Perturbation");
+%% E.1.1.1 restriction
+Uopt_lf = Uopt_hf(1:10:end, :); % pick every 10th elem
+% Uopt_lf = repelem(Uopt_lf, 10, 1); % uncomment to visualize
+% vis_sols(Uopt_hf, Uopt_lf, data.lqrsol{1}.times, "HF obj", perturb_range, "Perturbation");
 
[…]

### 2024-11-18 artefacts/paper.tex
derive analytical form for correlation

@@
 \usepackage{hyperref} % Required for hyperlinks
+\usepackage{amssymb} % Required for probability symbols
 
@@
 \section{Problem Formulation}
+This secition describes how the state space model of the dynamics is set up, along with the controller design and cost function for the trajectory optimization problem.
 \subsection{Mass-Spring-Damper System}
-For this paper, we consider a simplified trajectory optimization problem that is easy to understand and analyze.
-We try to solve a reference tracking problem for the position and velocity of a simple mass-spring-damper system.
+For this paper, a simplified trajectory optimization problem is considered that is easy to understand and analyze.
+The objective is to solve a reference tracking problem for the position and velocity of a simple mass-spring-damper system.
 The mass-spring-damper system has the following parameters:
@@
 
-We track the position $x$ and velocity $\dot{x}$ of the mass-spring-damper system using the control input $u$, which applies the force $F$ on the mass.
-These are the states and control inputs of the system, giving use the following dynamics
+The position $x$ and velocity $\dot{x}$ of the mass-spring-damper system can be changed using the control input $u$,
+which applies the force $F$ on the mass. This yields the following dynamics
 
@@
 
-For a closed-loop system, we need to observe the output $y$ of the system, and in this case we observe the position $x$.
+For a closed-loop system, the output $y$ of the system needs to be observed, which is both the position and velocity in this case.
 
@@
   \begin{bmatrix}
-    1 & 0
+    1 & 0 \\
+    0 & 1
   \end{bmatrix}
@@
 
-This gives us an LTI system in the form of $\dot{x} = Ax + Bu$ and $y = Cx$, where
+This yields an LTI system in the form of $\dot{x} = Ax + Bu$ and $y = Cx$, where
 
@@
     \begin{bmatrix}
-      1 & 0
+      1 & 0 \\
+      0 & 1
     \end{bmatrix}
@@
 
-\subsection{LQR Controller}
+\subsection{LQR Controller} \label{lqr_ctrl}
 We set up an LQR controller to track the reference position and velocity of the mass-spring-damper system.
@@
 
-\section{Control Variates}
+\subsection{Stochastic LQR Controller}
+We will consider the case where the initial state $x_0$ is stochastic, with mean $\mu$ and covariance $\Sigma$.
+The control input vector $U$ is still deterministic.
+
+In this case, the state vector $X$ is stochastic, owing to the the initial condition $x_0$ being a vector of random variables. The state propagation equations are now expectations
+
+\begin{equation}
+  \begin{aligned}
+    \mathbb{E}[x_1] &= \mathbb{E}[A x_0 + B u_0] = A \mathbb{E}[x_0] + B u_0 \\
+    \mathbb{E}[x_2] &= \mathbb{E}[A x_1 + B u_1] = A \mathbb{E}[x_1] + B u_1 = A^2 \mathbb{E}[x_0] + A B u_0 + B u_1 \\
+    \vdots \\
+    \mathbb{E}[x_N] &= \mathbb{E}[A x_{N-1} + B u_{N-1}] = A \mathbb{E}[x_{N-1}] + B u_{N-1} \\
+    &= A^N \mathbb{E}[x_0] + A^{N-1} B u_0 + A^{N-2} B u_1 + \ldots + B u_{N-1} \\
+  \end{aligned}
+\end{equation}
+
+The matrices $S$ and $M$ remain the same, and we have the following expression for the expected value of the state vector $X$
+
+$$
+\mathbb{E}[X] = S U + M \mathbb{E}[x_0]
+$$
+
+The cost matrices $\bar{Q}$ and $\bar{R}$ are also the same. The cost function $J$ is now a random variable,
+with an expected value $\mathbb{E}[J]$ and variance $Var[J]$.
+
+The expected value of the cost function over all possible realizations of the initial conditions is
+$$
+\begin{aligned}
+  \mathbb{E}[J] &= \mathbb{E}[x_N^T P x_N + \sum_{k=0}^{N-1} (x_k^T Q x_k + u_k^T R u_k)] \\
+  &= \mathbb{E}[X^T \bar{Q} X + U^T \bar{R} U + x_0^T Q x_0] \\
+  &= \mathbb{E}[X^T \bar{Q} X] + \mathbb{E}[U^T \bar{R} U] + \mathbb{E}[x_0^T Q x_0] \\
+  &= \mathbb{E}[(S U + M x_0)^T \bar{Q} (S U + M x_0)] + U^T \bar{R} U + \mathbb{E}[x_0^T Q x_0] \\
+  &= \mathbb{E}[U^T S^T \bar{Q} S U + U^T S^T \bar{Q} M x_0 + x_0^T M^T \bar{Q} S U + x_0^T M^T \bar{Q} M x_0] + U^T \bar{R} U + \mathbb{E}[x_0^T] Q \mathbb{E}[x_0] + \text{tr}(Q \text{Cov}[x_0]) \\
+  &= U^T S^T \bar{Q} S U + \mathbb{E}[U^T S^T \bar{Q} M x_0] + \mathbb{E}[x_0^T M^T \bar{Q} S U] + \mathbb{E}[x_0^T M^T \bar{Q} M x_0] + U^T \bar{R} U + \mathbb{E}[x_0^T] Q \mathbb{E}[x_0] + \text{tr}(Q \text{Cov}[x_0]) \\
+  &= U^T (S^T \bar{Q} S + \bar{R}) U + U^T S^T \bar{Q} M \mathbb{E}[x_0] + \mathbb{E}[x_0^T] M^T \bar{Q} S U + \mathbb{E}[x_0^T M^T \bar{Q} M x_0] + \mathbb{E}[x_0^T] Q \mathbb{E}[x_0] + \text{tr}(Q \text{Cov}[x_0]) \\
+  &= U^T (S^T \bar{Q} S + \bar{R}) U + \mathbb{E}[x_0^T] M^T \bar{Q} S U + \mathbb{E}[x_0^T] M^T \bar{Q} S U + \mathbb{E}[x_0^T] M^T \bar{Q} M \mathbb{E}[x_0] + \text{tr}(M^T \bar{Q} M \text{Cov}[x_0]) \hdots \\
+  & \quad + \mathbb{E}[x_0^T] Q \mathbb{E}[x_0] + \text{tr}(Q \text{Cov}[x_0]) \\
+  &= U^T (S^T \bar{Q} S + \bar{R}) U + 2 \mathbb{E}[x_0^T] M^T \bar{Q} S U + \mathbb{E}[x_0^T] (M^T \bar{Q} M + Q) \mathbb{E}[x_0] + \text{tr}((M^T \bar{Q} M + Q) \text{Cov}[x_0]) \\
+\end{aligned}
+$$
+
+Assuming $\mathbb{E}[x_0]$ and $\text{Cov}[x_0]$ is calculated using a Monte Carlo estimator, we can write the cost function as
+
+$$
+\begin{aligned}
+  H &= S^T \bar{Q} S + \bar{R} = H^T \\
+  q &= (\mathbb{E}[x_0^T] M^T \bar{Q} S)^T = S^T \bar{Q} M \mathbb{E}[x_0] \\
+  c &= \mathbb{E}[x_0^T] (M^T \bar{Q} M + Q) \mathbb{E}[x_0] + \text{tr}((M^T \bar{Q} M + Q) \text{Cov}[x_0])
+\end{aligned}
+$$
+
+The cost function can now be written as a quadratic function of the control inputs $U$
+
+$$
+J = U^T H U + 2 q^T U + c
+$$
+
+Which has a gradient with respect to $U$ as
+
+$$
+\Delta J = 2 H U + 2 q^T
+$$
+
+Assuming minimum exists, we can set the gradient to zero to get the optimal control input $U^*$
+
+$$
+\begin{aligned}
+  \Delta J &= 2 H U + 2 q^T = 0 \\
+  \implies U^* &= -H^{-1} q
+\end{aligned}
+$$
+
+% ref for quadratic form: https://en.wikipedia.org/wiki/Quadratic_form_(statistics)
+
+Tthe cost function can be rewritten as
+
+$$
+\begin{aligned}
+  J &= U^T(S^T \bar{Q} S + \bar{R}) U + 2 x_0^T M^T \bar{Q} S U + x_0^T(M^T \bar{Q} M + Q) x_0 \\
+  \text{Let}& \\
+  K &= U^T(S^T \bar{Q} S + \bar{R}) U \\
+  L &= 2 M^T \bar{Q} S U \\
+  N &= M^T \bar{Q} M + Q = N^T \\
+  &\text{(quadratic multiplication by diagnonal($\bar{Q}$) results in a symmetric, $Q$ is symmetric)} \\
+  \text{Then}& \\
+  J &= K + x_0^T L + x_0^T N x_0 \\
+\end{aligned}
+$$
+
+Then, the expectation and variance of the cost function is as follows. Take note that the random variable here is $x_0$ and $\mathbb{E}[x_0]$ is a scalar.
+
+% ref:
+% - https://www.math.uwaterloo.ca/~hwolkowi/matrixcookbook.pdf
+% - https://math.stackexchange.com/questions/2163694/expectation-of-quartic-form-for-multivariate-gaussian/4247240#4247240
+% - https://math.stackexchange.com/questions/1302882/variance-of-a-quadratic-form
+
+\begin{equation}
+  \begin{aligned}
+    \mathbb{E}[J] &= \mathbb{E}[K] + \mathbb{E}[x_0^T L] + \mathbb{E}[x_0^T N x_0] \\
+    &= K + \mathbb{E}[x_0^T] L + \mathbb{E}[x_0^T] N \mathbb{E}[x_0] + \text{tr}(N \text{Cov}[x_0]) \\
+    Var[J] &= Var[K + x_0^T L + x_0^T N x_0] \\
+    &\text{note: $K$ is deterministic, so its variance is 0} \\
+    & \text{note: $x_0^T L$ is a scalar, $x_0^T L = L^T x_0$} \\
+    &= Var[L^T x_0 + x_0^T N x_0] \\
+    &= 2\text{tr}(N \text{Cov}[x_0] N \text{Cov}[x_0]) + 4 \mathbb{E}[x_0^T]N\text{Cov}[x_0]N\mathbb{E}[x_0] + 4 L^T \text{Cov}[x_0] N \mathbb{E}[x_0] + L^T \text{Cov}[x_0] L \\
+  \end{aligned}
+\end{equation}
+
+To determine the covariance between the costs of two solutions $U_1$ and $U_2$, the components of $J$ that depend on $U$
+need to be written in terms of $U_1$ and $U_2$.
+
+For $U_1$, the cost function is
+
+$$
+\begin{aligned}
+  J(U_1) &= K_1 + x_0^T L_1 + x_0^T N x_0 \\
+  \text{where} \\
+  K_1 &= U_1^T(S^T \bar{Q} S + \bar{R}) U_1 \\
+  L_1 &= 2 M^T \bar{Q} S U_1 \\
+\end{aligned}
+$$
+
+Similarly, for $U_2$, the cost function is
+
+$$
+\begin{aligned}
+  J(U_2) &= K_2 + x_0^T L_2 + x_0^T N x_0 \\
+  \text{where} \\
+  K_2 &= U_2^T(S^T \bar{Q} S + \bar{R}) U_2 \\
+  L_2 &= 2 M^T \bar{Q} S U_2 \\
+\end{aligned}
+$$
+
+% ref:
+% - https://math.stackexchange.com/questions/1302882/variance-of-a-quadratic-form
+% - https://stats.stackexchange.com/questions/303466/prove-that-mathrmcovxtax-xtbx-2-mathrmtra-sigma-b-sigma-4-mu
+% - https://stats.stackexchange.com/questions/35084/covariance-of-a-linear-and-quadratic-form-of-a-multivariate-normal
+% - https://www.researchgate.net/publication/243770190_Quadratic_Forms_in_Random_Variables P74, 3.2d.9
+the covariance between their costs can be calculated as
+\begin{equation}
+  \begin{aligned}
+    \text{Cov}[J(U_1), J(U_2)] &= \text{Cov}[K_1 + x_0^T L_1 + x_0^T N x_0, K_2 + x_0^T L_2 + x_0^T N x_0] \\
+    &= \text{Cov}[x_0^T L_1 + x_0^T N x_0, x_0^T L_2 + x_0^T N x_0] \text{ (covariance is shift invariant)} \\
+    &= \text{Cov}[x_0^T L_1, x_0^T L_2] + \text{Cov}[x_0^T L_1, x_0^T N x_0] + \text{Cov}[x_0^T N x_0, x_0^T L_2] + \text{Cov}[x_0^T N x_0, x_0^T N x_0] \\
+    &= \text{Cov}[ L_1^T x_0, L_2^T x_0] + \text{Cov}[L_1^T x_0, x_0^T N x_0] + \text{Cov}[x_0^T N x_0, L_2^T x_0] + \text{Cov}[x_0^T N x_0, x_0^T N x_0] \\
+    &= L_1^T \text{Cov}[x_0] L_2 + \text{Cov}[x_0^T N x_0, L_1^T x_0] + \text{Cov}[x_0^T N x_0, x_0^T L_2] + \text{var}[x_0^T N x_0] \\
+    &= L_1^T \text{Cov}[x_0] L_2 + 2\mathbb{E}[x_0^T] N \text{Cov}[x_0] L_1 + 2\mathbb{E}[x_0^T] N \text{Cov}[x_0] L_2 \hdots \\
+    & \quad 4 \mathbb{E}[x_0^T]N\text{Cov}[x_0]N\mathbb{E}[x_0] + 2\text{tr}((N \text{Cov}[x_0])^2) \\
+  \end{aligned}
+\end{equation}
+
+\section{Multi Level Monte Carlo}
+
+This section formulates the stochastic trajectory optimization problem as a multi-level Monte Carlo problem.
 
@@
 After converting them to the same dimensions as done in section \ref{comp_opt},
-we can compute the cost for some realizations of the initial state $x_0$.
-This will give us a set of costs of both high and low fidelity control inputs for a certain model/cost function.
-We can then compute the correlation coefficient $\rho$ between the costs.
-To illustrate this, assume we have $N$ realizations of the initial state $x_0$, and J is the cost function.
+the cost for some realizations of the initial state $x_0$ can be computed.
+Hence, a set of costs of both high and low fidelity control inputs can be computed for a certain model/cost function.
+The correlation coefficient $\rho$ can then be computed between the distribution of the cost function values.
+To illustrate this, assume there are $N$ realizations of the initial state $x_0$, and $J$ is the cost function.
 $u_h$ and $u_l$ are the control inputs for the high and low fidelity models respectively, where the dimensions have been
@@
 \begin{equation}
-  \rho = \frac{1}{N-1} \sum_{i=1}^{N} \left( \frac{J(u_h, x_0^i) - \mu_h}{\sigma_h} \right) \left( \frac{J(u_l, x_0^i) - \mu_l}{\sigma_l} \right)
+  \rho(J(u_h), J(u_l)) = \frac{1}{N-1} \sum_{i=1}^{N} \left( \frac{J(u_h, x_0^i) - \mu_h}{\sigma_h} \right) \left( \frac{J(u_l, x_0^i) - \mu_l}{\sigma_l} \right)
+\end{equation}
+
+This can be simplified to:
+
+\begin{equation}
+  \rho(J(u_h), J(u_l)) = \frac{\text{Cov}(J(u_h), J(u_l))}{\sigma_h \sigma_l}
 \end{equation}
@@
 This correlation coefficient $\rho$ can be calculated using the MATLAB function \href{https://www.mathworks.com/help/matlab/ref/corrcoef.html#bunkaln}{\texttt{corrcoef}}.
+
+The correlation can also be calculated analytically using the covariance of the costs in the high fidelity objective function and the low fidelity objective function.
+Let
+
+\begin{itemize}
+  \item $J_h$: high fidelity cost function
+  \item $U_1$: high fidelity control input
+  \item $U_2$: upsampled low fidelity control input $u_{lh} = T_{lh} u_l$
+  \item $J_1 = J_h(U_1, x_0)$
+  \item $\sigma_1$: standard deviation of $J_1$, $\sqrt{\text{Var}[J_1]}$
+  \item $\sigma_2$: standard deviation of $J_2$, $\sqrt{\text{Var}[J_2]}$
+\end{itemize}
+
+Then correlation coefficient between $J_1$ and $J_2$ is
+
+\begin{equation}
+  \begin{aligned}
+    \rho(J_1, J_2) &= \frac{\text{Cov}(J_1, J_2)}{\sigma_1 \sigma_2} \\
+    &= \frac{\text{Cov}(J_1, J_2)}{\sqrt{\text{Var}[J_1]} \sqrt{\text{Var}[J_2]}}
+  \end{aligned}
+\end{equation}
+
+Where $\text{Cov}(J_1, J_2)$, $\text{Var}[J_1]$ and $\text{Var}[J_2]$ can be calculated using the eqations in section \ref{lqr_ctrl}.
 

### 2024-11-19 artefacts/paper.tex
add correct covariance calculation

@@
 
-Tthe cost function can be rewritten as
+The cost function can be rewritten as
 
@@
 
-To determine the covariance between the costs of two solutions $U_1$ and $U_2$, the components of $J$ that depend on $U$
-need to be written in terms of $U_1$ and $U_2$.
+To determine the covariance between the costs of two solutions $U_1$ and $U_2$, in different cost functions
+$J_1$ and $J_2$, we need to first rewrite the cost functions to differentiate between them.
 
@@
 \begin{aligned}
-  J(U_1) &= K_1 + x_0^T L_1 + x_0^T N x_0 \\
+  J_1(U_1) &= K_1 + x_0^T L_1 + x_0^T N_1 x_0 \\
   \text{where} \\
-  K_1 &= U_1^T(S^T \bar{Q} S + \bar{R}) U_1 \\
-  L_1 &= 2 M^T \bar{Q} S U_1 \\
+  K_1 &= U_1^T(S_1^T \bar{Q}_1 S_1 + \bar{R}_1) U_1 \\
+  L_1 &= 2 M_1^T \bar{Q}_1 S_1 U_1 \\
+  N_1 &= M_1^T \bar{Q}_1 M + Q_1 \\
+  \text{and} \\
+  & S_1, M_1, \bar{Q}_1, \bar{R}_1, Q_1 \text{ depend on the discrectization timestep of } J_1
 \end{aligned}
@@
 \begin{aligned}
-  J(U_2) &= K_2 + x_0^T L_2 + x_0^T N x_0 \\
+  J_2(U_2) &= K_2 + x_0^T L_2 + x_0^T N_2 x_0 \\
   \text{where} \\
-  K_2 &= U_2^T(S^T \bar{Q} S + \bar{R}) U_2 \\
-  L_2 &= 2 M^T \bar{Q} S U_2 \\
+  K_2 &= U_2^T(S_2^T \bar{Q}_2 S_2 + \bar{R}_2) U_2 \\
+  L_2 &= 2 M_2^T \bar{Q}_2 S_2 U_2 \\
+  N_2 &= M_2^T \bar{Q}_2 M + Q_2 \\
+  \text{and} \\
+  & S_2, M_2, \bar{Q}_2, \bar{R}_2, Q_2 \text{ depend on the discrectization timestep of } J_2
 \end{aligned}
@@
 % - https://stats.stackexchange.com/questions/35084/covariance-of-a-linear-and-quadratic-form-of-a-multivariate-normal
-% - https://www.researchgate.net/publication/243770190_Quadratic_Forms_in_Random_Variables P74, 3.2d.9
 the covariance between their costs can be calculated as
+
+\begin{equation}\label{eq:covariance}
+  \begin{aligned}
+    \text{Cov}[J_1(U_1), J_2(U_2)] &= \text{Cov}[K_1 + x_0^T L_1 + x_0^T N_1 x_0, K_2 + x_0^T L_2 + x_0^T N_2 x_0] \\
+    &= \text{Cov}[x_0^T L_1 + x_0^T N_1 x_0, x_0^T L_2 + x_0^T N_2 x_0] \text{ (covariance is shift invariant)} \\
+    &= \text{Cov}[x_0^T L_1, x_0^T L_2] + \text{Cov}[x_0^T L_1, x_0^T N_2 x_0] + \text{Cov}[x_0^T N_1 x_0, x_0^T L_2] + \text{Cov}[x_0^T N_1 x_0, x_0^T N_2 x_0] \\
+  \end{aligned}
+\end{equation}
+
+Determining the covariance of each component in equation \ref{eq:covariance},
+
 \begin{equation}
   \begin{aligned}
-    \text{Cov}[J(U_1), J(U_2)] &= \text{Cov}[K_1 + x_0^T L_1 + x_0^T N x_0, K_2 + x_0^T L_2 + x_0^T N x_0] \\
-    &= \text{Cov}[x_0^T L_1 + x_0^T N x_0, x_0^T L_2 + x_0^T N x_0] \text{ (covariance is shift invariant)} \\
-    &= \text{Cov}[x_0^T L_1, x_0^T L_2] + \text{Cov}[x_0^T L_1, x_0^T N x_0] + \text{Cov}[x_0^T N x_0, x_0^T L_2] + \text{Cov}[x_0^T N x_0, x_0^T N x_0] \\
-    &= \text{Cov}[ L_1^T x_0, L_2^T x_0] + \text{Cov}[L_1^T x_0, x_0^T N x_0] + \text{Cov}[x_0^T N x_0, L_2^T x_0] + \text{Cov}[x_0^T N x_0, x_0^T N x_0] \\
-    &= L_1^T \text{Cov}[x_0] L_2 + \text{Cov}[x_0^T N x_0, L_1^T x_0] + \text{Cov}[x_0^T N x_0, x_0^T L_2] + \text{var}[x_0^T N x_0] \\
-    &= L_1^T \text{Cov}[x_0] L_2 + 2\mathbb{E}[x_0^T] N \text{Cov}[x_0] L_1 + 2\mathbb{E}[x_0^T] N \text{Cov}[x_0] L_2 \hdots \\
-    & \quad 4 \mathbb{E}[x_0^T]N\text{Cov}[x_0]N\mathbb{E}[x_0] + 2\text{tr}((N \text{Cov}[x_0])^2) \\
+    \text{Cov}[x_0^T L_1, x_0^T L_2] &= \text{Cov}[L_1^T x_0, L_2^T x_0] \text{ transpose of a scalar is itself} \\
+    &= L_1^T \text{Cov}[x_0] L_2 \\
+  \end{aligned}
+\end{equation}
+
+\begin{equation}
+  % - https://www.researchgate.net/publication/243770190_Quadratic_Forms_in_Random_Variables P74, 3.2d.9
+  \begin{aligned}
+    \text{Cov}[x_0^T L_1, x_0^T N_2 x_0] &= \text{Cov}[x_0^T N_2 x_0, L_1^T x_0] \text{ transpose of a scalar is itself} \\
+    &= 2 \mathbb{E}[x_0^T] N_2 \text{Cov}[x_0] L_1 \\
+  \end{aligned}
+\end{equation}
+
+\begin{equation}
+  % - https://www.researchgate.net/publication/243770190_Quadratic_Forms_in_Random_Variables P74, 3.2d.9
+  \begin{aligned}
+    \text{Cov}[x_0^T N_1 x_0, x_0^T L_2] &= \text{Cov}[x_0^T N_1 x_0, L_2^T x_0] \text{ transpose of a scalar is itself} \\
+    &= 2 \mathbb{E}[x_0^T] N_1 \text{Cov}[x_0] L_2 \\
+  \end{aligned}
+\end{equation}
+
+\begin{equation}
+  % - https://www.researchgate.net/publication/243770190_Quadratic_Forms_in_Random_Variables P75, 3.2d.12
+  \begin{aligned}
+    \text{Cov}[x_0^T N_1 x_0, x_0^T N_2 x_0] &= 2 \text{tr}(N_1 \text{Cov}[x_0] N_2 \text{Cov}[x_0]) + 4 \mathbb{E}[x_0^T]N_1\text{Cov}[x_0]N_2\mathbb{E}[x_0] \\
+  \end{aligned}
+\end{equation}
+
+Combining these, we get the covariance as
+
+\begin{equation}
+  \begin{aligned}
+    \text{Cov}[J_1(U_1), J_2(U_2)] &= L_1^T \text{Cov}[x_0] L_2 + 2 \mathbb{E}[x_0^T] N_2 \text{Cov}[x_0] L_1 + 2 \mathbb{E}[x_0^T] N_1 \text{Cov}[x_0] L_2 \hdots \\
+    & \quad + 2 \text{tr}(N_1 \text{Cov}[x_0] N_2 \text{Cov}[x_0]) + 4 \mathbb{E}[x_0^T]N_1\text{Cov}[x_0]N_2\mathbb{E}[x_0]
   \end{aligned}
@@
   \item $J_h$: high fidelity cost function
+  \item $J_l$: low fidelity cost function
   \item $U_1$: high fidelity control input
-  \item $U_2$: upsampled low fidelity control input $u_{lh} = T_{lh} u_l$
+  \item $U_2$: downsampled high fidelity control input $= T_{hlr} U_1$ or $T_{hla} U_1$
   \item $J_1 = J_h(U_1, x_0)$
-  \item $\sigma_1$: standard deviation of $J_1$, $\sqrt{\text{Var}[J_1]}$
-  \item $\sigma_2$: standard deviation of $J_2$, $\sqrt{\text{Var}[J_2]}$
+  \item $J_2 = J_l(U_2, x_0)$
+  \item $\sigma_1$: standard deviation of $J_1$, $=\sqrt{\text{Var}[J_1]}$
+  \item $\sigma_2$: standard deviation of $J_2$, $=\sqrt{\text{Var}[J_2]}$
 \end{itemize}
@@
 
-Where $\text{Cov}(J_1, J_2)$, $\text{Var}[J_1]$ and $\text{Var}[J_2]$ can be calculated using the eqations in section \ref{lqr_ctrl}.
+Where $\text{Cov}(J_1, J_2)$, $\text{Var}[J_1]$ and $\text{Var}[J_2]$ can be calculated using the equations in section \ref{lqr_ctrl}.
 

### 2024-11-19 tests.m

@@ -92,3 +92,24 @@ term = x' * A + diag(x' * B * x);
 st_var = var(term);
 an_var = 2 * trace(B * x_cov * B * x_cov) + A' * x_cov * A + 4 * (x_mean' * B + A') * x_cov * B * x_mean;
 fprintf("Analytic: %f, Statistic: %f\n", an_var, st_var);
+
+% covariance
+disp(newline + "Covariance");
+K1 = randn;
+K2 = randn;
+L1 = randn(nx, 1);
+L2 = randn(nx, 1);
+N1 = randn(nx, nx);
+N1 = N1 * N1';
+N2 = randn(nx, nx);
+N2 = N2 * N2';
+term1 = zeros(1, samples);
+term2 = zeros(1, samples);
+for i = 1:samples
+	term1(i) = K1 + x(:, i)' * L1 + x(:, i)' * N1 * x(:, i);
+	term2(i) = K2 + x(:, i)' * L2 + x(:, i)' * N2 * x(:, i);
+end
+st_cov = cov(term1', term2'); % cov needs row vectors
+st_cov = st_cov(1, 2); % only need the off diagonal element
+an_cov = L1'*x_cov*L2 + 2*x_mean'*N2*x_cov*L1 + 2*x_mean'*N1*x_cov*L2 + 2*trace(N1*x_cov*N2*x_cov) + 4*x_mean'*N1*x_cov*N2*x_mean;
+fprintf("Analytic: %f, Statistic: %f\n", an_cov, st_cov);

### 2024-11-19 stoch_traj.m
consolidate cost calculation for all samples for correlation, log graphs

@@ -213,24 +213,26 @@ Uopt_lf = Uopt_hf(1:10:end, :); % pick every 10th elem
 % vis_sols(Uopt_hf, Uopt_lf, data.lqrsol{1}.times, "HF obj", perturb_range, "Perturbation");
 
 % calculate costs and correlation for HF/LF sols for each perturbation
-[cost_hf, cost_lf] = calc_costs_multifid(data.lqrsol{1}, data.lqrsol{2}, Uopt_hf, Uopt_lf);
-corr = calc_corr_multifid(x0_rv, u0, ref, data.lqrsol{1}, data.lqrsol{2}, Uopt_hf, Uopt_lf);
-% plot costs
+[cost_hf, cost_lf] = calc_costs_multifid(x0_rv, u0, ref, data.lqrsol{1}, data.lqrsol{2}, Uopt_hf, Uopt_lf);
+corr = calc_corr_multifid(cost_hf, cost_lf);
+% [cov_st, cov_an] = calc_cov(cost_hf, cost_lf, x0_rv, u0, ref, data.lqrsol{1}, data.lqrsol{2}, Uopt_hf, Uopt_lf);
+%% plot costs
+% plot_cov(cov_st, cov_an);
 title_str = ["$J_h$ and $J_l$", "$J_h$ (restriction)"];
 costs_str = ["$J_h(u_h)$", "$J_l(u_{hlr})$"];
-plot_multifid_costs(perturb_range, cost_hf, cost_lf, corr, title_str, costs_str, "Perturbation");
+plot_multifid_costs(perturb_range, mean(cost_hf,2), mean(cost_lf,2), corr, title_str, costs_str, "Perturbation");
 
 %% E.1.1.2 averaging
 Uopt_lf = downsample_avg(Uopt_hf, 10); % get HF in LF by avg every 10 values of HF
 % vis_sols(Uopt_hf, Uopt_lf, data.lqrsol{1}.times, "HF obj", perturb_range, "Perturbation");
 Uopt_lf = Uopt_lf(1:10:end, :); % pick every 10th elem
 % calculate costs and correlation for HF/LF sols for each perturbation
-[cost_hf, cost_lf] = calc_costs_multifid(data.lqrsol{1}, data.lqrsol{2}, Uopt_hf, Uopt_lf);
[…]

### 2024-11-19 stoch_traj.m
add covariance calculation and plotting

@@ -209,17 +209,17 @@ Uopt_hf = data.lqrsol{1}.Uopt + perturbation; % perturb hf sol
 
 %% E.1.1.1 restriction
 Uopt_lf = Uopt_hf(1:10:end, :); % pick every 10th elem
-% Uopt_lf = repelem(Uopt_lf, 10, 1); % uncomment to visualize
-% vis_sols(Uopt_hf, Uopt_lf, data.lqrsol{1}.times, "HF obj", perturb_range, "Perturbation");
+% Uopt_lf_vis = repelem(Uopt_lf, 10, 1); % uncomment to visualize
+% vis_sols(Uopt_hf, Uopt_lf_vis, data.lqrsol{1}.times, "HF obj", perturb_range, "Perturbation");
 
 % calculate costs and correlation for HF/LF sols for each perturbation
 [cost_hf, cost_lf] = calc_costs_multifid(x0_rv, u0, ref, data.lqrsol{1}, data.lqrsol{2}, Uopt_hf, Uopt_lf);
 corr = calc_corr_multifid(cost_hf, cost_lf);
-% [cov_st, cov_an] = calc_cov(cost_hf, cost_lf, x0_rv, u0, ref, data.lqrsol{1}, data.lqrsol{2}, Uopt_hf, Uopt_lf);
+[cov_st, cov_an] = calc_cov(cost_hf, cost_lf, x0_rv, u0, ref, data.lqrsol{1}, data.lqrsol{2}, Uopt_hf, Uopt_lf);
 %% plot costs
-% plot_cov(cov_st, cov_an);
 title_str = ["$J_h$ and $J_l$", "$J_h$ (restriction)"];
 costs_str = ["$J_h(u_h)$", "$J_l(u_{hlr})$"];
+plot_cov(perturb_range, cov_st, cov_an, title_str, "Perturbation");
 plot_multifid_costs(perturb_range, mean(cost_hf,2), mean(cost_lf,2), corr, title_str, costs_str, "Perturbation");
 
 %% E.1.1.2 averaging
@@ -229,9 +229,11 @@ Uopt_lf = Uopt_lf(1:10:end, :); % pick every 10th elem
 % calculate costs and correlation for HF/LF sols for each perturbation
 [cost_hf, cost_lf] = calc_costs_multifid(x0_rv, u0, ref, data.lqrsol{1}, data.lqrsol{2}, Uopt_hf, Uopt_lf);
[…]

### 2024-11-20 stoch_traj.m
add analytical correlation calculation and graphs

@@ -43,7 +43,7 @@ R = eye(nu) * 10; % input cost
 % we dont want to penalize the state being away from the origin
 P = eye(nx) * 0;
 
-rv_samples = 1000;
+rv_samples = 10000;
 perturb_dir_samples = 1000;
 perturb_dir_mag = 0.1;
 perturb_range = -1:0.25:6;
@@ -61,7 +61,7 @@ ylabel('Position [m]/Velocity [m/s]');
 ylim(plot_ylim_l);
 title("Open loop dynamics");
 xlabel('Time [s]');
-legend show; legend boxoff;
+legend show;
 grid on;
 saveas(fig, 'figs/ol_det.svg');
 
@@ -95,7 +95,7 @@ ylabel('Position [m]');
 ylim(plot_ylim_l);
 title("Open loop dynamics with stochastic initial state (" + rv_samples + " samples)");
 xlabel('Time [s]');
-legend show; legend boxoff;
+legend show;
 grid on;
 saveas(fig, 'figs/ol_stoch_init.svg');
 
@@ -138,7 +138,7 @@ for idx = 1:length(TsList)
   ax.YColor = 'r';
   title("(Ts:"+Ts+") Closed loop dynamics");
   xlabel('Time [s]');
-  legend show; legend boxoff;
+  legend show;
   grid on;
   saveas(fig, "figs/"+Ts+"_cl_det.svg");
   
@@ -175,7 +175,7 @@ for idx = 1:length(TsList)
   ax.YColor = 'r';
   title("(Ts:"+Ts+") Closed loop dynamics with stochastic initial state (" + rv_samples + " samples)");
   xlabel('Time [s]');
-  legend show; legend boxoff;
+  legend show;
   grid on;
   saveas(fig, "figs/"+Ts+"_cl_stoch_init.svg");
   
@@ -195,7 +195,7 @@ for idx = 1:length(TsList)
   ylabel('Probability Density');
   title("(Ts:"+Ts+") Cost distribution of stochastic LQR ("+rv_samples+" samples)");
   xlabel('Cost');
-  legend show; legend boxoff;
+  legend show;
   grid on;
   saveas(fig, "figs/"+Ts+"_cost_dist.svg");
 end
[…]

### 2024-11-20 stoch_traj.m
optimize code with parfor and vectorization of plots and cost calculation

@@ -43,7 +43,7 @@ R = eye(nu) * 10; % input cost
 % we dont want to penalize the state being away from the origin
 P = eye(nx) * 0;
 
-rv_samples = 10000;
+rv_samples = 1000;
 perturb_dir_samples = 1000;
 perturb_dir_mag = 0.1;
 perturb_range = -1:0.25:6;
@@ -75,20 +75,17 @@ x0_rv_mean_ext = [x0_rv_mean; u0; ref];
 x0_rv_cov_ext = blkdiag(x0_rv_cov, zeros(3,3));
 
 %% C.1 open loop dynamics
-data.x_ol_stoch = zeros(nx, Tsim/Tfid + 1, rv_samples);
-for i = 1:rv_samples
+x_ol_stoch = zeros(nx, Tsim/Tfid + 1, rv_samples);
+parfor i = 1:rv_samples
   [~, x_ode] = ode45(@(t, x) msd(t, x, 0, A_msd, B_msd), 0:Tfid:Tsim, x0_rv(:, i));
-  data.x_ol_stoch(:, :, i) = x_ode';
+  x_ol_stoch(:, :, i) = x_ode';
 end
+data.x_ol_stoch = x_ol_stoch;
 
 %% C.2 plot open loop dynamics
 fig = figure;
 hold on;
-
-for i = 1:rv_samples
-  plot(data.t, data.x_ol_stoch(1, :, i), 'b', 'LineWidth', 1, 'HandleVisibility', 'off', 'Color', [0.5 0.5 0.5, 0.2]);
-end
-
+plot(data.t, squeeze(data.x_ol_stoch(1, :, :)), 'b', 'LineWidth', 1, 'HandleVisibility', 'off', 'Color', [0.5 0.5 0.5, 0.2]);
 plot(data.t, data.x_ol(1, :), 'b', 'LineWidth', 2, 'DisplayName', 'Position: Deterministic', 'Color', [0 0 1, 0.5]);
 plot(data.t, mean(data.x_ol_stoch(1, :, :), 3), '--k', 'LineWidth', 2, 'DisplayName', 'Position: Sample Average');
 ylabel('Position [m]');
@@ -146,23 +143,19 @@ for idx = 1:length(TsList)
   % Uopt only depends on x0, so we can use the same Kopt
   Uopt = Kopt * x0_rv_mean_ext;
[…]

### 2024-12-12 artefacts/paper.tex
plot convergence distance from optimum for mlmc

@@
 
-Where the number of columns after which the next one is selected is determined by the ratio $r_{hl} = n_{u_l} / n_{u_h}$.
+Where the number of columns after which the next one is selected is determined by the ratio $r_{hl} = n_{u_h} / n_{u_l}$.
 This gives us $u_{hlr} = T_{hlr} u_h$, which can be compared with $u_l$.
 
-To downsample $u_h$ to $u_l$ by averaging, we need the ratio $r_{lh} = n_{u_h} \ n_{u_l}$.
+To downsample $u_h$ to $u_l$ by averaging, we need the ratio $r_{lh} = n_{u_h} / n_{u_l}$.
 We can then multiply $u_h$ with a matrix $T_{hla}$ of size $n_{u_l} \times n_{u_h}$:

### 2024-12-12 stoch_traj.m

@@ -245,9 +245,10 @@ plot_multifid_costs(perturb_range, mean(cost_hf,2), mean(cost_lf,2), corr_st, co
 %% *** E.2 correlation at different points in a numerical optimizer
 %% E.2.1 num opt with HF objective fn
 Uopt_hf = data.lqrsol{1}.Uopt;
-% u0_num = zeros(size(Uopt_hf));
 u0_num = repelem(data.lqrsol{2}.Uopt, 10, 1); % warm start
-global num_opt_iters num_opt_data;
+% u0_num = zeros(size(Uopt_hf)); % zero start
+% u0_num = randn(size(Uopt_hf)); % random start
+global num_opt_iters num_opt_data num_opt_fvals;
 num_opt_iters = 0;
 num_opt_data = zeros(size(Uopt_hf,1), 100);
 fun = @(u) LQRCostwithGrad(data.lqrsol{1}.x0_ext, u, data.lqrsol{1}.Q_ext, data.lqrsol{1}.S, data.lqrsol{1}.M, data.lqrsol{1}.Qbar, data.lqrsol{1}.Rbar);
@@ -256,7 +257,7 @@ Uopt_num = fminunc(fun, u0_num, options);
 uopt_diff = sqrt(sum((Uopt_hf - Uopt_num).^2));
 
 U_hf = num_opt_data(:, 1:num_opt_iters);
-
+U_num_hf = U_hf; % save for plotting later
 % this is just to get the costs for the averaged HF solution in LF sim as the MLMC estimator would
 %% E.2.1.1 correlation with restriction
 U_lf = U_hf(1:10:end, :);
@@ -300,8 +301,9 @@ fun = @(u) calc_mlmc_est(data.lqrsol{1}, data.lqrsol{2}, u, cv_samples, cv_sampl
 options = optimoptions('fminunc', 'Display', 'iter', 'OutputFcn', @outfun);
 Uopt_num = fminunc(fun, u0_num, options);
 U_hf = num_opt_data(:, 1:num_opt_iters);
+U_num_mlmc = U_hf; % save for plotting later
 U_lf = downsample_avg(U_hf, 10);
-vis_sols(U_hf, U_lf, data.lqrsol{1}.times, "Solutions along optimizer path", 1:num_opt_iters, "Iteration");
[…]

### 2024-12-12 stoch_traj.m
use actual mean and cov in mlmc

@@ -297,8 +297,8 @@ num_opt_iters = 0;
 num_opt_data = zeros(size(Uopt_hf,1), 100);
 global alpha_mc;
 alpha_mc = zeros(1, 100);
-fun = @(u) calc_mlmc_est(data.lqrsol{1}, data.lqrsol{2}, u, cv_samples, cv_samples, u0, ref, x0_rv);
-options = optimoptions('fminunc', 'Display', 'iter', 'OutputFcn', @outfun);
+fun = @(u) calc_mlmc_est(data.lqrsol{1}, data.lqrsol{2}, u, cv_samples, cv_samples, u0, ref, x0_rv, x0_mean, x0_cov);
+options = optimoptions('fminunc', 'Display', 'iter', 'OutputFcn', @outfun); % finite diff
 Uopt_num = fminunc(fun, u0_num, options);
 U_hf = num_opt_data(:, 1:num_opt_iters);
 U_num_mlmc = U_hf; % save for plotting later
@@ -330,8 +330,8 @@ u_lf = U_lf(:, it_max_cor);
 num_opt_iters = 0;
 num_opt_data = zeros(size(Uopt_hf,1), 100);
 alpha_mc = zeros(1, 100);
-fun = @(u) calc_mlmc_est_max_corr(data.lqrsol{1}, data.lqrsol{2}, u, u_lf, cv_samples, cv_samples, u0, ref, x0_rv);
-options = optimoptions('fminunc', 'Display', 'iter', 'OutputFcn', @outfun);
+fun = @(u) calc_mlmc_est_max_corr(data.lqrsol{1}, data.lqrsol{2}, u, u_lf, cv_samples, cv_samples, u0, ref, x0_rv, x0_mean, x0_cov);
+options = optimoptions('fminunc', 'Display', 'iter', 'OutputFcn', @outfun); % finite dif
 Uopt_num = fminunc(fun, u0_num, options);
 U_hf = num_opt_data(:, 1:num_opt_iters);
 U_num_mlmc_fix = U_hf; % save for plotting later
@@ -358,7 +358,7 @@ ylabel("Alpha");
 grid on;
 
 %% plot convergence distance comparison
[…]

### 2024-12-12 stoch_traj.m
calculate estimator variance for all estimators

@@ -361,95 +361,132 @@ grid on;
 plot_convergence_dist(data.lqrsol{1}.Uopt, U_num_hf, U_num_mlmc, U_num_mlmc_fix, ["HF", "CV", "CV with max corr"], "Convergence distance comparion");
 
 %% G Convergence and variance with various optimizers and sample sizes
-% num_rv_samples = [10, 100, 500, 1000];
-% num_estimator_samples = 100;
-% u0_num = repelem(data.lqrsol{2}.Uopt, 10, 1); % warm start
-% max_iters = 20;
-% options = optimoptions('fminunc', 'Display', 'iter', 'OutputFcn', @outfun, "MaxIterations", max_iters);
-% for num_samples=num_rv_samples
-%   fprintf("\n*** RV Samples: %d ***\n", num_samples);
-%   temp_cost.hf = zeros(num_estimator_samples, max_iters);
-%   temp_cost.mlmc = zeros(num_estimator_samples, max_iters);
-%   temp_cost.mlmc_fix = zeros(num_estimator_samples, max_iters);
-%   for i=1:num_estimator_samples
-%     % get the RV samples
-%     x0_rv = mvnrnd(x0_mean, x0_cov, num_samples)';
-%     x0_rv_ext = [x0_rv; repmat(u0, 1, num_samples); repmat(ref, 1, num_samples)];
+num_rv_samples = [10, 100, 500, 1000];
+num_estimator_samples = 10;
+u0_num = repelem(data.lqrsol{2}.Uopt, 10, 1); % warm start
+max_iters = 30;
+options = optimoptions('fminunc', 'Display', 'iter', 'OutputFcn', @outfun, "MaxIterations", max_iters);
+data.var_hf = zeros(length(num_rv_samples), max_iters);
+data.var_mlmc = zeros(length(num_rv_samples), max_iters);
+data.var_mlmc_fix = zeros(length(num_rv_samples), max_iters);
+for num_samples=num_rv_samples
+  fprintf("\n*** RV Samples: %d ***\n", num_samples);
+  wait_bar = waitbar(0, "Running estimators with " + num_samples + " samples");
+  temp_cost.hf = zeros(num_estimator_samples, max_iters);
+  temp_cost.mlmc = zeros(num_estimator_samples, max_iters);
[…]

### 2024-12-16 stoch_traj.m
clculate and plot convergence

@@ -362,13 +362,14 @@ plot_convergence_dist(data.lqrsol{1}.Uopt, U_num_hf, U_num_mlmc, U_num_mlmc_fix,
 
 %% G Convergence and variance with various optimizers and sample sizes
 num_rv_samples = [10, 100, 500, 1000];
-num_estimator_samples = 10;
+num_estimator_samples = 100;
 u0_num = repelem(data.lqrsol{2}.Uopt, 10, 1); % warm start
 max_iters = 30;
 options = optimoptions('fminunc', 'Display', 'iter', 'OutputFcn', @outfun, "MaxIterations", max_iters);
-data.var_hf = zeros(length(num_rv_samples), max_iters);
-data.var_mlmc = zeros(length(num_rv_samples), max_iters);
-data.var_mlmc_fix = zeros(length(num_rv_samples), max_iters);
+data.hf_var = zeros(length(num_rv_samples), max_iters);
+data.mlmc_var = zeros(length(num_rv_samples), max_iters);
+data.mlmc_fix_var = zeros(length(num_rv_samples), max_iters);
+data.hf_u = zeros(length(u0_num), max_iters, num_estimator_samples, length(num_rv_samples));
 for num_samples=num_rv_samples
   fprintf("\n*** RV Samples: %d ***\n", num_samples);
   wait_bar = waitbar(0, "Running estimators with " + num_samples + " samples");
@@ -388,7 +389,8 @@ for num_samples=num_rv_samples
     x0_rv_ext_mean = mean(x0_rv_ext, 2);
     fun = @(u) mean(LQRCost_vec(x0_rv_ext, u, data.lqrsol{1}.Q_ext, data.lqrsol{1}.S, data.lqrsol{1}.M, data.lqrsol{1}.Qbar, data.lqrsol{1}.Rbar));
     Uopt_num = fminunc(fun, u0_num, options);
-    temp_cost.hf(i, :) = num_opt_fvals(end-max_iters+1:end);
+    temp_cost.hf(i, :) = num_opt_fvals(1:max_iters);
+    data.hf_u(:, :, i, num_rv_samples == num_samples) = num_opt_data(:, 1:max_iters);
     
     % for MLMC estimator
     num_opt_iters = 0;
@@ -396,7 +398,8 @@ for num_samples=num_rv_samples
     num_opt_fvals = zeros(1, max_iters);
[…]

### 2024-12-17 stoch_traj.m
get analytical var over iterations with avg u

@@ -1,3 +1,14 @@
+% TODOS:
+% - we are assuming u0 to be 0, might not be true in general so might need to add u0 to cumsum
+% - extend MLMC to work for any number of levels
+% - make indices consistent when storing U in row vs column (make everything row based)
+% - split out all the stats related functions into a separate file
+% - calculate correlation in cv estimator as it goes along instead of all at once
+% - calculate 2d correlation analytically
+% - make U and u consistent in naming
+% - make usage of x_ext vs x, u and ref consistent, create ext extrenally and all functions should use that directly
+% - LQRcost_cov should take in the mean directly
+
 clc; clear; close all force;
 %% A. state space model
 
@@ -358,10 +369,10 @@ ylabel("Alpha");
 grid on;
 
 %% plot convergence distance comparison
-plot_convergence_dist(data.lqrsol{1}.Uopt, U_num_hf, U_num_mlmc, U_num_mlmc_fix, ["HF", "CV", "CV with max corr"], "Convergence distance comparion");
+plot_convergence_dist(data.lqrsol{1}.Uopt, U_num_hf, U_num_mlmc, U_num_mlmc_fix, ["HF", "CV", "CV with max corr"], "Convergence distance comparison");
 
 %% G Convergence and variance with various optimizers and sample sizes
-num_rv_samples = [10, 100, 500, 1000];
+num_rv_samples = [100];
 num_estimator_samples = 100;
 u0_num = repelem(data.lqrsol{2}.Uopt, 10, 1); % warm start
 max_iters = 30;
@@ -370,6 +381,8 @@ data.hf_var = zeros(length(num_rv_samples), max_iters);
 data.mlmc_var = zeros(length(num_rv_samples), max_iters);
 data.mlmc_fix_var = zeros(length(num_rv_samples), max_iters);
 data.hf_u = zeros(length(u0_num), max_iters, num_estimator_samples, length(num_rv_samples));
[…]

### 2024-12-18 artefacts/paper.tex
plot analytical and statistical variance

@@
 \begin{itemize}
-  \item Mass: $m = 5$ g
-  \item Spring constant: $k = 2$ N/m
-  \item Damping coefficient: $c = 0.5$ Ns/m
+  \item $m = 5$ g : Mass
+  \item $k = 2$ N/m : Spring constant
+  \item $c = 0.5$ Ns/m : Damping coefficient
     % \item Initial position: $x_0 = 0$ m
@@
 
-\subsection{Stochastic LQR Controller}
+\subsection{Stochastic LQR Controller} \label{stochastic_lqr}
 We will consider the case where the initial state $x_0$ is stochastic, with mean $\mu$ and covariance $\Sigma$.
@@
 
-\section{Multi Level Monte Carlo}
+\section{Monte Carlo with Control Variates}
 
-This section formulates the stochastic trajectory optimization problem as a multi-level Monte Carlo problem.
+This section formulates the stochastic trajectory optimization problem as a Monte Carlo problem with a low fidelity control variate.
 
@@
 
-To determine if the low fidelity model can be used as a control variate, we need to determine if it coreelates with the high fidelity model.
+To determine if the low fidelity model can be used as a control variate, we need to determine if it correlates with the high fidelity model.
 
@@
 \begin{itemize}
-  \item High fidelity model/cost function: $J_h(u, x_0)$
-  \item Low fidelity model/cost function: $J_l(u, x_0)$
-  \item Control input in high fidelity model: $u_h$
-  \item Size of $u_h$: $n_{u_h} \times 1$
-  \item Control input in low fidelity model: $u_l$
-  \item Size of $u_l$: $n_{u_l} \times 1$
-  \item Random initial state: $x_0$
-  \item High fidelity optimum: $u_h^*$
-  \item Low fidelity optimum: $u_l^*$
+  \item $J_h(u, x_0)$: High fidelity model/cost function
+  \item $J_l(u, x_0)$: Low fidelity model/cost function
+  \item $u_h$: Control input in high fidelity model, $n_{u_h} \times 1$
+  \item $u_l$: Control input in low fidelity model, $n_{u_l} \times 1$
+  \item $x_0$: Random initial state
+  \item $u_h^*$: High fidelity optimum
+  \item $u_l^*$: Low fidelity optimum
 \end{itemize}
@@
 \begin{itemize}
-  \item High fidelity model/cost function: $J_h$
-  \item Low fidelity model/cost function: $J_l$
-  \item High fidelity optimum: $u_h^*$
-  \item High fidelity perturbation: $p_h$
-  \item Perturbed high fidelity solution: $u_h = u_h^* + p_h$
-  \item Perturbed high fidelity solution downsampled using restriction: $u_{hlr} = T_{hlr} u_h$
-  \item Perturbed high fidelity solution downsampled using averaging: $u_{hla} = T_{hla} u_h$
+  \item $J_h$: High fidelity model/cost function
+  \item $J_l$: Low fidelity model/cost function
+  \item $u_h^*$: High fidelity optimum
+  \item $p_h$: High fidelity perturbation
+  \item $u_h = u_h^* + p_h$: Perturbed high fidelity solution
+  \item $u_{hlr} = T_{hlr} u_h$: Perturbed high fidelity solution downsampled using restriction:
+  \item $u_{hla} = T_{hla} u_h$: Perturbed high fidelity solution downsampled using averaging
 \end{itemize}
@@
 \begin{itemize}
-  \item High fidelity model/cost function: $J_h$
-  \item Low fidelity model/cost function: $J_l$
-  \item Low fidelity optimum: $u_l^*$
-  \item Low fidelity perturbation: $p_l$
-  \item Perturbed low fidelity solution: $u_l = u_l^* + p_l$
-  \item Perturbed high fidelity solution upsampled: $u_{lh} = T_{lh} u_l$
+  \item $J_h$: High fidelity model/cost function
+  \item $J_l$: Low fidelity model/cost function
+  \item $u_l^*$: Low fidelity optimum
+  \item $p_l$: Low fidelity perturbation
+  \item $u_l = u_l^* + p_l$: Perturbed low fidelity solution
+  \item $u_{lh} = T_{lh} u_l$: Perturbed high fidelity solution upsampled
 \end{itemize}
@@
 \begin{itemize}
-  \item High fidelity model/cost function: $J_h$
-  \item Low fidelity model/cost function: $J_l$
-  \item High fidelity solution: $u_h$
-  \item High fidelity solution downsampled using restriction: $u_{hlr} = T_{hlr} u_h$
-  \item High fidelity solution downsampled using averaging: $u_{hla} = T_{hla} u_h$
+  \item $J_h$: High fidelity model/cost function
+  \item $J_l$: Low fidelity model/cost function
+  \item $u_h$: High fidelity solution
+  \item $u_{hlr} = T_{hlr} u_h$: High fidelity solution downsampled using restriction
+  \item $u_{hla} = T_{hla} u_h$: High fidelity solution downsampled using averaging
 \end{itemize}
@@
 
+It is also possible to compute the correlation between $J_h(u_h, x_0)$ and $J_l(u_{hlr}, x_0)$ for every iteration of the numerical optimizer.
+This helps us determine if there is a certain iteration of $J_l(u_{hlr}, x_0)$ that correlates well with all iterations of $J_h(u_h, x_0)$.
+The same comparison can be made with $J_l(u_{hla}, x_0)$.
+
+Comparing correlation between $J_h(u_h, x_0)$ and $J_h(u_{hlr}, x_0)$ across all iterations
+
+Comparing correlation between $J_h(u_h, x_0)$ and $J_l(u_{hla}, x_0)$ across all iterations
+
+\subsection{Control Variate Construction}
+
+The control variate is constructed using the low fidelity model/cost function, since we have found that there is some correlation between $J_h(u_h, x_0)$ and $J_l(u_{hla}, x_0)$.
+Here we are only using the averaged downsampled high fidelity solution $u_{hla}$ as the control variate.
+The important variables for constructing the control variate are:
+
+\begin{itemize}
+  \item $x_0$: Random initial state
+  \item $J_h$: High fidelity model/cost function
+  \item $J_l$: Low fidelity model/cost function
+  \item $u_h$: High fidelity solution
+  \item $u_{hla} = T_{hla} u_h$: High fidelity solution downsampled using averaging
+  \item $S_n^{MC}(J_h(u_h, x_0))$: Monte Carlo estimate of the high fidelity cost function
+  \item $S_n^{MC}(J_l(u_{hla}, x_0))$: Monte Carlo estimate of the low fidelity cost function
+  \item $\alpha$: Control variate coefficient
+\end{itemize}
+
+The control variate coefficient $\alpha$ can be calculated as
+
+\begin{equation}
+  \alpha = -\frac{\text{Cov}(J_h(u_h, x_0), J_l(u_{hla}, x_0))}{\text{Var}[J_l(u_{hla}, x_0)]}
+\end{equation}
+
+Hence, the control variate is
+
+\begin{equation}
+  S_n^{CV}(J_h, J_l) = S_n^{MC}(J_h(u_h, x_0)) + \alpha (S_n^{MC}(J_l(u_{hla}, x_0)) - \mathbb{E}[J_l(u_{hla}, x_0)])
+\end{equation}
+
+Where the samples of $x_0$ used in $J_h(u_h, x_0)$ and $J_l(u_{hla}, x_0)$ are the same.
+
+Note that $\text{Cov}(J_h(u_h, x_0), J_l(u_{hla}, x_0))$, $\text{Var}[J_l(u_{hla}, x_0)]$ and $\mathbb{E}[J_l(u_{hla}, x_0)]$ are calculated analytically using the equations in section \ref{stochastic_lqr}.
+
+\subsection{Numerical optimizer using Control Variate}
+
+The numerical optimizer was run in 3 different scenarios
+
+\begin{itemize}
+  \item $S_n^{MC}(J_h(u_h, x_0))$: MC estimate of the high fidelity cost function
+  \item $S_n^{CV}(J_h, J_l)$: CV estimate with $u_{hla}$ at the current iteration
+  \item $S_n^{CV_{max}}(J_h, J_l)$:CV estimate with $u_{hla}$ at the iteration with the highest correlation
+\end{itemize}
+
+For these experiments, $n$, the number of samples of $x_0$ used in the MC estimate, was compared between 10, 100, 500 and 1000.
+The comparison is done between $S_n^{MC}(J_h(u_h, x_0))$ vs $S_{\frac{n}{2}}^{CV}(J_h, J_l)$ and $S_{\frac{n}{2}}^{CV_{max}}(J_h, J_l)$ to keep the number of cost function evaluations the same with the control variate estimators having a lower computational cost due to half the sammples being used in the low fidelity cost function.
+
+To compare the variance, we ran a numerical optimizer with n random samples of $x_0$ and calculated the variance for each iteration.
+
+The variance was also calculated analytically for each control variate
 \end{document}

### 2024-12-18 stoch_traj.m

@@ -372,23 +372,20 @@ grid on;
 plot_convergence_dist(data.lqrsol{1}.Uopt, U_num_hf, U_num_mlmc, U_num_mlmc_fix, ["HF", "CV", "CV with max corr"], "Convergence distance comparison");
 
 %% G Convergence and variance with various optimizers and sample sizes
-num_rv_samples = [100];
+num_rv_samples = [10 100 1000];
 num_estimator_samples = 100;
 u0_num = repelem(data.lqrsol{2}.Uopt, 10, 1); % warm start
 max_iters = 30;
-options = optimoptions('fminunc', 'Display', 'iter', 'OutputFcn', @outfun, "MaxIterations", max_iters);
-data.hf_var = zeros(length(num_rv_samples), max_iters);
-data.mlmc_var = zeros(length(num_rv_samples), max_iters);
-data.mlmc_fix_var = zeros(length(num_rv_samples), max_iters);
+options = optimoptions('fminunc', 'Display', 'iter', 'OutputFcn', @outfun, "MaxIterations", max_iters, "OptimalityTolerance", 1e-12);
+data.hf_cost = zeros(max_iters, num_estimator_samples, length(num_rv_samples));
+data.mlmc_cost = zeros(max_iters, num_estimator_samples, length(num_rv_samples));
+data.mlmc_fix_cost = zeros(max_iters, num_estimator_samples, length(num_rv_samples));
 data.hf_u = zeros(length(u0_num), max_iters, num_estimator_samples, length(num_rv_samples));
 data.mlmc_u = zeros(length(u0_num), max_iters, num_estimator_samples, length(num_rv_samples));
 data.mlmc_fix_u = zeros(length(u0_num), max_iters, num_estimator_samples, length(num_rv_samples));
 for num_samples=num_rv_samples
   fprintf("\n*** RV Samples: %d ***\n", num_samples);
   wait_bar = waitbar(0, "Running estimators with " + num_samples + " samples");
-  temp_cost.hf = zeros(num_estimator_samples, max_iters);
-  temp_cost.mlmc = zeros(num_estimator_samples, max_iters);
[…]

### 2025-02-17 artefacts/paper.tex
clean up write-up till control variate/monte carlo

@@
 \usepackage{amssymb} % Required for probability symbols
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 
@@
 \subsection{Mass-Spring-Damper System}
-For this paper, a simplified trajectory optimization problem is considered that is easy to understand and analyze.
+For this paper, a simplified trajectory optimization problem is considered for ease of modelling.
 The objective is to solve a reference tracking problem for the position and velocity of a simple mass-spring-damper system.
 The mass-spring-damper system has the following parameters:
+
 \begin{itemize}
@@
 
-\begin{equation}
-  m\ddot{x} = F - c\dot{x} - kx
-  \implies \ddot{x} = \frac{F}{m} - \frac{c}{m}\dot{x} - \frac{k}{m}x
-\end{equation}
+$$
+\begin{aligned}
+  m\ddot{x} &= F - c\dot{x} - kx \\
+  \implies \ddot{x} &= \frac{F}{m} - \frac{c}{m}\dot{x} - \frac{k}{m}x
+\end{aligned}
+$$
 
-The position $x$ and velocity $\dot{x}$ of the mass-spring-damper system can be changed using the control input $u$,
-which applies the force $F$ on the mass. This yields the following dynamics
+We have the state as $[x \dot{x}]^T$ containing the position and velocity of the mass, both of which are observed.
+The control input $u=F$ is applied to the mass, yielding the following state space model:
 
-\begin{equation}
+$$
+\begin{aligned}
   \begin{bmatrix}
@@
     \ddot{x}
-  \end{bmatrix} =
+  \end{bmatrix} &=
   \begin{bmatrix}
@@
     \frac{1}{m}
-  \end{bmatrix} u
-\end{equation}
-
-For a closed-loop system, the output $y$ of the system needs to be observed, which is both the position and velocity in this case.
-
-\begin{equation}
-  y =
+  \end{bmatrix} u \\
+  y &=
   \begin{bmatrix}
@@
   \end{bmatrix}
-\end{equation}
+\end{aligned}
+$$
 
-This yields an LTI system in the form of $\dot{x} = Ax + Bu$ and $y = Cx$, where
+Thus, we have a standard continuous-time LTI system $\dot{x} = A_cx + B_cu$ and $y = C_cx$, where
 
@@
   \begin{aligned}
-    A &=
+    A_c &=
     \begin{bmatrix}
@@
     \end{bmatrix} \\
-    B &=
+    B_c &=
     \begin{bmatrix}
@@
     \end{bmatrix} \\
-    C &=
+    C_c &=
     \begin{bmatrix}
@@
 \subsection{LQR Controller} \label{lqr_ctrl}
+The continuous-time system can be converted to discrete time using such that $x_{k+1} = A x_k + B u_k$ and $y_k = C x_k$ for a chosen sampling time $T_s$.
+The $A$ and $B$ matrices for the discrete time system are then used to formulate the cost function for the LQR problem.
+
 We set up an LQR controller to track the reference position and velocity of the mass-spring-damper system.
@@
 
-To set up the LQR controller, we need to define the cost function of the system, which consists of:
+The LQR controller minimizes the cost function
+
+$$
+\min_{u_0, u_1 ... u_{N-1}}J = \sum_{k=0}^{N-1} (e_k^T Q_e e_k + \Delta u_k^T R \Delta u_k)
+$$
+
+Where
 
 \begin{itemize}
-  \item $u_0, u_1 ... u_{N-1}$ : control inputs at each timestep of the trajectory
+  \item $\Delta u_0, \Delta u_1 ... \Delta u_{N-1}$ : control inputs at each timestep of the trajectory
   \item $J$ : total cost to be minimized
@@
 \end{itemize}
-The LQR controller minimizes the cost function
 
-\begin{equation}
-  \min_{u_0, u_1 ... u_{N-1}}J = \sum_{k=0}^{N-1} (e_k^T Q_e e_k + \Delta u_k^T R \Delta u_k)
-\end{equation}
+Note that we can add a terminal penalty using the DARE solution for the LQR problem, but for simplicity we will not consider it here.
 
-\subsection{Stochastic LQR Controller} \label{stochastic_lqr}
-We will consider the case where the initial state $x_0$ is stochastic, with mean $\mu$ and covariance $\Sigma$.
-The control input vector $U$ is still deterministic.
+To track the error $e_k$, we extend the state vector $x_k$ to include the control input $u_{k-1}$ and the reference output $r_k$.
 
-In this case, the state vector $X$ is stochastic, owing to the the initial condition $x_0$ being a vector of random variables. The state propagation equations are now expectations
+$$
+x_k^{ext} =
+\begin{bmatrix}
+  x_k \\
+  u_{k-1} \\
+  r_k
+\end{bmatrix}
+$$
+
+With the resulting state space model
 
@@
   \begin{aligned}
-    \mathbb{E}[x_1] &= \mathbb{E}[A x_0 + B u_0] = A \mathbb{E}[x_0] + B u_0 \\
-    \mathbb{E}[x_2] &= \mathbb{E}[A x_1 + B u_1] = A \mathbb{E}[x_1] + B u_1 = A^2 \mathbb{E}[x_0] + A B u_0 + B u_1 \\
-    \vdots \\
-    \mathbb{E}[x_N] &= \mathbb{E}[A x_{N-1} + B u_{N-1}] = A \mathbb{E}[x_{N-1}] + B u_{N-1} \\
-    &= A^N \mathbb{E}[x_0] + A^{N-1} B u_0 + A^{N-2} B u_1 + \ldots + B u_{N-1} \\
+    x_{k+1}^{ext} &=
+    \begin{bmatrix}
+      A & B & 0 \\
+      0 & \mathbb{I}_{n_u \times n_u} & 0 \\
+      0 & 0 & \mathbb{I}_{n_r \times n_r}
+    \end{bmatrix}
+    \begin{bmatrix}
+      x_k \\
+      u_{k-1} \\
+      r_k
+    \end{bmatrix} +
+    \begin{bmatrix}
+      B \\
+      \mathbb{I}_{n_u \times n_u} \\
+      0
+    \end{bmatrix} \Delta u_k \\
   \end{aligned}
@@
 
-The matrices $S$ and $M$ remain the same, and we have the following expression for the expected value of the state vector $X$
+with
 
 $$
-\mathbb{E}[X] = S U + M \mathbb{E}[x_0]
-$$
-
-The cost matrices $\bar{Q}$ and $\bar{R}$ are also the same. The cost function $J$ is now a random variable,
-with an expected value $\mathbb{E}[J]$ and variance $Var[J]$.
-
-The expected value of the cost function over all possible realizations of the initial conditions is
-$$
 \begin{aligned}
-  \mathbb{E}[J] &= \mathbb{E}[x_N^T P x_N + \sum_{k=0}^{N-1} (x_k^T Q x_k + u_k^T R u_k)] \\
-  &= \mathbb{E}[X^T \bar{Q} X + U^T \bar{R} U + x_0^T Q x_0] \\
-  &= \mathbb{E}[X^T \bar{Q} X] + \mathbb{E}[U^T \bar{R} U] + \mathbb{E}[x_0^T Q x_0] \\
-  &= \mathbb{E}[(S U + M x_0)^T \bar{Q} (S U + M x_0)] + U^T \bar{R} U + \mathbb{E}[x_0^T Q x_0] \\
-  &= \mathbb{E}[U^T S^T \bar{Q} S U + U^T S^T \bar{Q} M x_0 + x_0^T M^T \bar{Q} S U + x_0^T M^T \bar{Q} M x_0] + U^T \bar{R} U + \mathbb{E}[x_0^T] Q \mathbb{E}[x_0] + \text{tr}(Q \text{Cov}[x_0]) \\
-  &= U^T S^T \bar{Q} S U + \mathbb{E}[U^T S^T \bar{Q} M x_0] + \mathbb{E}[x_0^T M^T \bar{Q} S U] + \mathbb{E}[x_0^T M^T \bar{Q} M x_0] + U^T \bar{R} U + \mathbb{E}[x_0^T] Q \mathbb{E}[x_0] + \text{tr}(Q \text{Cov}[x_0]) \\
-  &= U^T (S^T \bar{Q} S + \bar{R}) U + U^T S^T \bar{Q} M \mathbb{E}[x_0] + \mathbb{E}[x_0^T] M^T \bar{Q} S U + \mathbb{E}[x_0^T M^T \bar{Q} M x_0] + \mathbb{E}[x_0^T] Q \mathbb{E}[x_0] + \text{tr}(Q \text{Cov}[x_0]) \\
-  &= U^T (S^T \bar{Q} S + \bar{R}) U + \mathbb{E}[x_0^T] M^T \bar{Q} S U + \mathbb{E}[x_0^T] M^T \bar{Q} S U + \mathbb{E}[x_0^T] M^T \bar{Q} M \mathbb{E}[x_0] + \text{tr}(M^T \bar{Q} M \text{Cov}[x_0]) \hdots \\
-  & \quad + \mathbb{E}[x_0^T] Q \mathbb{E}[x_0] + \text{tr}(Q \text{Cov}[x_0]) \\
-  &= U^T (S^T \bar{Q} S + \bar{R}) U + 2 \mathbb{E}[x_0^T] M^T \bar{Q} S U + \mathbb{E}[x_0^T] (M^T \bar{Q} M + Q) \mathbb{E}[x_0] + \text{tr}((M^T \bar{Q} M + Q) \text{Cov}[x_0]) \\
+  A^{ext} &=
+  \begin{bmatrix}
+    A & B & 0 \\
+    0 & \mathbb{I}_{n_u \times n_u} & 0 \\
+    0 & 0 & \mathbb{I}_{n_r \times n_r}
+  \end{bmatrix} \\
+  B^{ext} &=
+  \begin{bmatrix}
+    B \\
+    \mathbb{I}_{n_u \times n_u} \\
+    0
+  \end{bmatrix}
 \end{aligned}
@@
 
-Assuming $\mathbb{E}[x_0]$ and $\text{Cov}[x_0]$ is calculated using a Monte Carlo estimator, we can write the cost function as
+Similary, $Qe$ with the extended state $x_k^{ext}$ transforms to $Q$ such that $e_k^T Q_e e_k = x_k^{ext^T} Q x_k^{ext}$.
+
+$$
+\begin{aligned}
+  e_k^T Q_e e_k &= (y_k - r_k)^T Q_e (y_k - r_k) \\
+  &= \left(
+    \begin{bmatrix} C & 0 & -\mathbb{I}_{n_r \times n_r}
+    \end{bmatrix}
+    \begin{bmatrix}
+      x_k \\
+      u_{k-1} \\
+      r_k
+  \end{bmatrix}\right)^T Q_e
+  \left(
+    \begin{bmatrix} C & 0 & -\mathbb{I}_{n_r \times n_r}
+    \end{bmatrix}
+    \begin{bmatrix}
+      x_k \\
+      u_{k-1} \\
+      r_k
+  \end{bmatrix}\right) \\
+  &= x_k^{ext^T}
+  \begin{bmatrix} C & 0 & -\mathbb{I}_{n_r \times n_r}
+  \end{bmatrix}^T Q_e
+  \begin{bmatrix} C & 0 & -\mathbb{I}_{n_r \times n_r}
+  \end{bmatrix} x_k^{ext} \\
+\end{aligned}
+$$
+
+\begin{equation}
+  Q =
+  \begin{bmatrix} C & 0 & -\mathbb{I}_{n_r \times n_r}
+  \end{bmatrix}^T Q_e
+  \begin{bmatrix} C & 0 & -\mathbb{I}_{n_r \times n_r}
+  \end{bmatrix}
+\end{equation}
+
+For simplicity, from now on we refer to $x_k^{ext}$ as $x_k$, $A^{ext}$ as $A$ and $B^{ext}$ as $B$.
+The cost function can now be written as
+
+$$
+J = \sum_{k=0}^{N-1} x_k^T Q x_k + \Delta u_k^T R \Delta u_k
+$$
+
+Using the state transition formula
+
+$$
+\begin{aligned}
+  x_1 &= A x_0 + B u_0 \\
+  x_2 &= A x_1 + B u_1 = A(A x_0 + B u_0) + B u_1 = A^2 x_0 + A B u_0 + B u_1 \\
+  \vdots \\
+  x_N &= A x_{N-1} + B u_{N-1} = A^{N-1} x_0 + A^{N-2} B u_0 + A^{N-3} B u_1 + \ldots + B u_{N-1} \\
+  \implies x_k &= A^k x_0 + \sum_{i=0}^{k-1} A^{k-1-i} B u_i
+\end{aligned}
+$$
+
+We construct the matrices $X$, $U$, $S$ and $M$ such that $X = SU + Mx_0$.
+
+$$
+\begin{aligned}
+  X &=
+  \begin{bmatrix}
+    x_1^{ext} \\
+    x_2^{ext} \\
+    \vdots \\
+    x_N^{ext}
+  \end{bmatrix} & \in \mathbb{R}^{Nn_x \times 1} \\
+  U &=
+  \begin{bmatrix}
+    \Delta u_0 \\
+    \Delta u_1 \\
+    \vdots \\
+    \Delta u_{N-1}
+  \end{bmatrix} & \in \mathbb{R}^{Nn_u \times 1} \\
+  S &=
+  \begin{bmatrix}
+    B & 0 & \ldots & 0 \\
+    AB & B & \ldots & 0 \\
+    \vdots & \vdots & \ddots & \vdots \\
+    A^{N-1} B & A^{N-2} B & \ldots & B
+  \end{bmatrix} & \in \mathbb{R}^{Nn_x \times Nn_u} \\
+  M &=
+  \begin{bmatrix}
+    A \\
+    A^2 \\
+    \vdots \\
+    A^N
+  \end{bmatrix} & \in \mathbb{R}^{Nn_x \times n_x}
+\end{aligned}
+$$
+
+Similarly, the matrices $\bar{Q}$ and $\bar{R}$ are constructed
+
+$$
+\begin{aligned}
+  \bar{Q} &=
+  \begin{bmatrix}
+    Q & 0 & \ldots & 0 \\
+    0 & Q & \ldots & 0 \\
+    \vdots & \vdots & \ddots & \vdots \\
+    0 & 0 & \ldots & 0
+  \end{bmatrix} \text{with size } Nn_x \times Nn_x \\
+  \bar{R} &=
+  \begin{bmatrix}
+    R & 0 & \ldots & 0 \\
+    0 & R & \ldots & 0 \\
+    \vdots & \vdots & \ddots & \vdots \\
+    0 & 0 & \ldots & R
+  \end{bmatrix} \text{with size } Nn_u \times Nn_u
+\end{aligned}
+$$
+
+The cost function can then be written as a quadratic function of the control inputs $U$
+
+\begin{equation}
+  \begin{aligned}
+    J &= \sum_{k=0}^{N-1} x_k^T Q x_k + \Delta u_k^T R \Delta u_k \\
+    % \text{ (extra $x_0$ term because it is not in $X$)} \\
+    &= X^T \bar{Q} X + U^T \bar{R} U + x_0^T Q x_0 \\
+    &= (S U + M x_0)^T \bar{Q} (S U + M x_0) + U^T \bar{R} U + x_0^T Q x_0 \\
+    % & \text{note } (A + B)^T = A^T + B^T \text{ and } (AB)^T = B^T A^T \\
+    &= (U^T S^T + x_0^T M^T) \bar{Q} (S U + M x_0) + U^T \bar{R} U + x_0^T Q x_0 \\
+    % & \text{note } A(B + C) = AB + AC \\
+    &= (U^T S^T + x_0^T M^T) (\bar{Q} S U + \bar{Q} M x_0) + U^T \bar{R} U + x_0^T Q x_0 \\
+    &= U^T S^T \bar{Q} S U + U^T S^T \bar{Q} M x_0 + x_0^T M^T \bar{Q} S U + x_0^T M^T \bar{Q} M x_0 + U^T \bar{R} U + x_0^T Q x_0 \\
+    % & \text{note } \bar{Q} \text{ is symmetric so it doesnt need to be transposed} \\
+    &= U^T S^T \bar{Q} S U + (x_0^T M^T \bar{Q} S U)^T + x_0^T M^T \bar{Q} S U + x_0^T M^T \bar{Q} M x_0 + U^T \bar{R} U + x_0^T Q x_0 \\
+    % & \text{note } x_0^T M^T \bar{Q} S U \text{ is a scalar(check its dimensions), so its transpose is itself} \\
+    &= U^T S^T \bar{Q} S U + x_0^T M^T \bar{Q} S U + x_0^T M^T \bar{Q} S U + x_0^T M^T \bar{Q} M x_0 + U^T \bar{R} U + x_0^T Q x_0 \\
+    % & \text{combining all quadratic and repeated terms} \\
+    &= U^T (S^T \bar{Q} S + \bar{R}) U + 2 x_0^T M^T \bar{Q} S U + x_0^T(M^T \bar{Q} M + Q) x_0\\
+  \end{aligned}
+\end{equation}
+
+Let
 
@@
   H &= S^T \bar{Q} S + \bar{R} = H^T \\
-  q &= (\mathbb{E}[x_0^T] M^T \bar{Q} S)^T = S^T \bar{Q} M \mathbb{E}[x_0] \\
-  c &= \mathbb{E}[x_0^T] (M^T \bar{Q} M + Q) \mathbb{E}[x_0] + \text{tr}((M^T \bar{Q} M + Q) \text{Cov}[x_0])
+  &\text{(quadratic multiplication by diagnonal($\bar{Q}$) results in a symmetric, R is symmetric)} \\
+  q &= (x_0^T M^T \bar{Q} S)^T = S^T \bar{Q} M x_0 \\
+  c &= x_0^T (M^T \bar{Q} M + Q) x_0
 \end{aligned}
@@
 $$
-\Delta J = 2 H U + 2 q^T
+\frac{\partial J}{\partial U} = 2 H U + 2 q^T
 $$
@@
 
+\begin{equation}
+  \begin{aligned}
+    \frac{\partial J}{\partial U} &= 2 H U + 2 q^T = 0 \\
+    \implies U^* &= -H^{-1} q
+  \end{aligned}
+\end{equation}
+
+\subsection{Stochastic LQR Controller} \label{stochastic_lqr}
+
+We consider the case where the initial state $x_0$ is stochastic, with mean $\mu$ and covariance $\Sigma$.
+Hence, the state vector $X$ is stochastic, since it is propagating the randomness of $x_0$ forward.
+The control input vector $U$ is deterministic. The cost function $J$ is now a random variable,
+with an expected value $\mathbb{E}[J]$ and variance $Var[J]$.
+
+\subsubsection{Solution of the Stochastic LQR controller}
+
+The expected value of the cost function over all possible realizations of $x_0$ is
+
+\begin{equation}
+  \begin{aligned}
+    \mathbb{E}[J] &= \mathbb{E}[\sum_{k=0}^{N-1} x_k^T Q x_k + \Delta u_k^T R \Delta u_k] \\
+    &= \mathbb{E}[X^T \bar{Q} X + U^T \bar{R} U + x_0^T Q x_0] \\
+    &= \mathbb{E}[X^T \bar{Q} X] + \mathbb{E}[U^T \bar{R} U] + \mathbb{E}[x_0^T Q x_0] \\
+    &= \mathbb{E}[(S U + M x_0)^T \bar{Q} (S U + M x_0)] + U^T \bar{R} U + \mathbb{E}[x_0^T Q x_0] \\
+    &= \mathbb{E}[U^T S^T \bar{Q} S U + U^T S^T \bar{Q} M x_0 + x_0^T M^T \bar{Q} S U + x_0^T M^T \bar{Q} M x_0] + U^T \bar{R} U + \mathbb{E}[x_0^T] Q \mathbb{E}[x_0] + \text{tr}(Q \text{Cov}[x_0]) \\
+    &= U^T S^T \bar{Q} S U + \mathbb{E}[U^T S^T \bar{Q} M x_0] + \mathbb{E}[x_0^T M^T \bar{Q} S U] + \mathbb{E}[x_0^T M^T \bar{Q} M x_0] + U^T \bar{R} U + \mathbb{E}[x_0^T] Q \mathbb{E}[x_0] + \text{tr}(Q \text{Cov}[x_0]) \\
+    &= U^T (S^T \bar{Q} S + \bar{R}) U + U^T S^T \bar{Q} M \mathbb{E}[x_0] + \mathbb{E}[x_0^T] M^T \bar{Q} S U + \mathbb{E}[x_0^T M^T \bar{Q} M x_0] + \mathbb{E}[x_0^T] Q \mathbb{E}[x_0] + \text{tr}(Q \text{Cov}[x_0]) \\
+    &= U^T (S^T \bar{Q} S + \bar{R}) U + \mathbb{E}[x_0^T] M^T \bar{Q} S U + \mathbb{E}[x_0^T] M^T \bar{Q} S U + \mathbb{E}[x_0^T] M^T \bar{Q} M \mathbb{E}[x_0] + \text{tr}(M^T \bar{Q} M \text{Cov}[x_0]) \dots \\
+    & \quad + \mathbb{E}[x_0^T] Q \mathbb{E}[x_0] + \text{tr}(Q \text{Cov}[x_0]) \\
+    &= U^T (S^T \bar{Q} S + \bar{R}) U + 2 \mathbb{E}[x_0^T] M^T \bar{Q} S U + \mathbb{E}[x_0^T] (M^T \bar{Q} M + Q) \mathbb{E}[x_0] + \text{tr}((M^T \bar{Q} M + Q) \text{Cov}[x_0]) \\
+  \end{aligned}
+\end{equation}
+
+Let
+
 $$
 \begin{aligned}
-  \Delta J &= 2 H U + 2 q^T = 0 \\
-  \implies U^* &= -H^{-1} q
+  H &= S^T \bar{Q} S + \bar{R} = H^T \\
+  q &= (\mathbb{E}[x_0^T] M^T \bar{Q} S)^T = S^T \bar{Q} M \mathbb{E}[x_0] \\
+  c &= \mathbb{E}[x_0^T] (M^T \bar{Q} M + Q) \mathbb{E}[x_0] + \text{tr}((M^T \bar{Q} M + Q) \text{Cov}[x_0])
 \end{aligned}
@@
 
-% ref for quadratic form: https://en.wikipedia.org/wiki/Quadratic_form_(statistics)
-
-The cost function can be rewritten as
+The expectation of cost function can be written as
 
 $$
-\begin{aligned}
-  J &= U^T(S^T \bar{Q} S + \bar{R}) U + 2 x_0^T M^T \bar{Q} S U + x_0^T(M^T \bar{Q} M + Q) x_0 \\
-  \text{Let}& \\
-  K &= U^T(S^T \bar{Q} S + \bar{R}) U \\
-  L &= 2 M^T \bar{Q} S U \\
-  N &= M^T \bar{Q} M + Q = N^T \\
-  &\text{(quadratic multiplication by diagnonal($\bar{Q}$) results in a symmetric, $Q$ is symmetric)} \\
-  \text{Then}& \\
-  J &= K + x_0^T L + x_0^T N x_0 \\
-\end{aligned}
+\mathbb{E}[J] = U^T H U + 2 q^T U + c
 $$
 
-Then, the expectation and variance of the cost function is as follows. Take note that the random variable here is $x_0$ and $\mathbb{E}[x_0]$ is a scalar.
+Which has a gradient with respect to $U$ as
 
+$$
+\frac{\partial \mathbb{E}[J]}{\partial U} = 2 H U + 2 q^T
+$$
+
+Assuming minimum exists, we can set the gradient to zero to get the optimal control input $U^*$ that minimizes the expected value of the cost function
+
+\begin{equation}
+  \begin{aligned}
+    \frac{\partial \mathbb{E}[J]}{\partial U} &= 2 H U + 2 q^T = 0 \\
+    \implies U^* &= -H^{-1} q
+  \end{aligned}
+\end{equation}
+
+\subsubsection{Variance and Covariance of cost function in terms of initial state}
+
+% ref for quadratic form: https://en.wikipedia.org/wiki/Quadratic_form_(statistics)
 % ref:
@@
 
+The expected value of the cost function can be rewritten as
+
+$$
+\mathbb{E}[J] = K + \mathbb{E}[x_0^T] L + \mathbb{E}[x_0^T] N \mathbb{E}[x_0] + \text{tr}(N \text{Cov}[x_0]) \\
+$$
+
+Where
+
+$$
+\begin{aligned}
+  x_0 &\sim \mathcal{N}(\mu, \Sigma) \\
+  K &= U^T(S^T \bar{Q} S + \bar{R}) U \\
+  L &= 2 M^T \bar{Q} S U \\
+  N &= M^T \bar{Q} M + Q = N^T \\
+  &\text{(quadratic multiplication by diagnonal($\bar{Q}$) results in a symmetric, $Q$ is symmetric)} \\
+\end{aligned}
+$$
+
+Using this, the variance can be calculated.
+
 \begin{equation}
   \begin{aligned}
-    \mathbb{E}[J] &= \mathbb{E}[K] + \mathbb{E}[x_0^T L] + \mathbb{E}[x_0^T N x_0] \\
-    &= K + \mathbb{E}[x_0^T] L + \mathbb{E}[x_0^T] N \mathbb{E}[x_0] + \text{tr}(N \text{Cov}[x_0]) \\
     Var[J] &= Var[K + x_0^T L + x_0^T N x_0] \\
@@
 
-\begin{equation}
+$$
 \begin{aligned}
@@
 \end{aligned}
-\end{equation}
+$$
 
-\begin{equation}
+$$
 % - https://www.researchgate.net/publication/243770190_Quadratic_Forms_in_Random_Variables P74, 3.2d.9
@@
 \end{aligned}
-\end{equation}
+$$
 
-\begin{equation}
+$$
 % - https://www.researchgate.net/publication/243770190_Quadratic_Forms_in_Random_Variables P75, 3.2d.12
@@
 \end{aligned}
-\end{equation}
+$$
 
@@
 \section{Monte Carlo with Control Variates}
+TODO: explain why we need control variates
 
@@
 \begin{itemize}
-  \item $S_n^{MC}(J_h(u_h, x_0))$: MC estimate of the high fidelity cost function
-  \item $S_n^{CV}(J_h, J_l)$: CV estimate with $u_{hla}$ at the current iteration
-  \item $S_n^{CV_{max}}(J_h, J_l)$:CV estimate with $u_{hla}$ at the iteration with the highest correlation
+  \item $S_n^{MC}(J_h(U_h, x_0))$: MC estimate of the high fidelity cost function
+  \item $S_n^{CV}(J_h(U_h, x_0), J_l(U_{hla}, x_0))$: CV estimate with $u_{hla}$ at the current iteration
+  \item $S_n^{CV_{max}}(J_h(U_h, x_0), J_l(U_{hla}^{max}, x_0))$:CV estimate where $u_{hla}^{max}$ is at the iteration with the highest correlation.
+\end{itemize}
+
+This is how the whole routine works, step-by-step:
+
+Initialization:
+\begin{itemize}
+  \item $U_l^* \gets$ Analytical solution of low fidelity model
+  \item $U_{h_0} \gets T_{lh} U_l^*$
+\end{itemize}
+
+% For $S_n^{CV}$ the numerical optimizer is run for $i^{max}$ iterations, where in each iteration $i$, the objective is calculated:
+% \begin{equation}
+%   S_n^{CV}(J_h, J_l) = S_n^{MC}(J_h(U, x_0)) + \alpha (S_n^{MC}(J_l(U_{hla}, x_0)) - \mathbb{E}[J_l(U_{hla}, x_0)])
+% \end{equation}
+
+% Where
+% \begin{itemize}
+%   \item $n$: Number of samples of $x_0$ used in the MC estimate to equal the cost of $S^{MC}(J_h(U, x_0))$ with all samples
+%   \item $U_{hla} = T_{hla} U$
+%   \item $\alpha = -\frac{\text{Cov}(J_h(U, x_0), J_l(U_{hla}, x_0))}{\text{Var}[J_l(U_{hla}, x_0)]}$ is calculated analytically
+%   \item $\mathbb{E}[J_l(U_{hla}, x_0)]$ is calculated analytically
+% \end{itemize}
+
+% For $S_n^{CV_{max}}$ the numerical optimizer is run for $i^{max}$ iterations, where in each iteration $i$, the objective is calculated:
+% \begin{equation}
+%   S_n^{CV_{max}}(J_h, J_l) = S_n^{MC}(J_h(U, x_0)) + \alpha (S_n^{MC}(J_l(U_{hla}^{max}, x_0)) - \mathbb{E}[J_l(U_{hla}^{max}, x_0)])
+% \end{equation}
+
+% Where
+% \begin{itemize}
+%   \item $n$: Number of samples of $x_0$ used in the MC estimate to equal the cost of $S^{MC}(J_h(U, x_0))$ with all samples
+%   \item $U_{hla} = T_{hla} U$
+%   \item $U_{hla}^{max} = \underset{U_{hla}(j)}{max} \{\rho(J_h(U(i), x_0^n), J_l(U_{hla}(j), x_0^n))\}_{j=1}^i$
+%   \item $\alpha = -\frac{\text{Cov}(J_h(U, x_0), J_l(U_{hla}^{max}, x_0))}{\text{Var}[J_l(U_{hla}^{max}, x_0)]}$ is calculated analytically
+%   \item $\mathbb{E}[J_l(U_{hla}^{max}, x_0)]$ is calculated analytically
+% \end{itemize}
+
+For $S_{nm}^{ACV}$ the numerical optimizer is run for $i^{max}$ iterations, where in each iteration $i$, the objective is calculated:
+\begin{equation}
+  S_{nm}^{ACV}(J_h, J_l) = S_n^{MC}(J_h(U, x_0)) + \alpha (S_n^{MC}(J_l(U_{hla}, x_0)) - S_m^{MC}(J_l(U_{hla}, x_0)))
+\end{equation}
+
+Where
+\begin{itemize}
+  \item $m$: Total number of samples of $x_0$
+  \item $n$: Number of samples of $x_0$ used in the MC estimate to equal the cost of $S_m^{MC}(J_h(U, x_0))$ (without considering the cost of $S_m^{MC}(J_l(U_{hla}, x_0))$)
+  \item $U_{hla} = T_{hla} U$
+  \item $\alpha = -\frac{\text{Cov}(J_h(U, x_0^n), J_l(U_{hla}, x_0^n))}{\text{Var}[J_l(U_{hla}, x_0^n)]}$ is calculated statistically with $n$ samples
+\end{itemize}
+
+For $S_{nm}^{ACV_{max}}$ the numerical optimizer is run for $i^{max}$ iterations, where in each iteration $i$, the objective is calculated:
+\begin{equation}
+  S_{nm}^{ACV_{max}}(J_h, J_l) = S_n^{MC}(J_h(U, x_0)) + \alpha (S_n^{MC}(J_l(U_{hla}^{max}, x_0)) - S_m^{MC}(J_l(U_{hla}^{max}, x_0)))
+\end{equation}
+
+Where
+\begin{itemize}
+  \item $m$: Total number of samples of $x_0$
+  \item $n$: Number of samples of $x_0$ used in the MC estimate to equal the cost of $S_m^{MC}(J_h(U, x_0))$ (without considering the cost of $S_m^{MC}(J_l(U_{hla}, x_0))$)
+  \item $U_{hla} = T_{hla} U$
+  \item $U_{hla}^{max} = \underset{U_{hla}(j)}{max} \{\rho(J_h(U(i), x_0^n), J_l(U_{hla}(j), x_0^n))\}_{j=1}^i$
+  \item $\alpha = -\frac{\text{Cov}(J_h(U, x_0^n), J_l(U_{hla}^{max}, x_0^n))}{\text{Var}[J_l(U_{hla}^{max}, x_0^n)]}$ is calculated statistically with $n$ samples
 \end{itemize}

### 2025-02-18 artefacts/paper.tex
add algorithm for CV and ACV

@@
 \section{Problem Formulation}
-This secition describes how the state space model of the dynamics is set up, along with the controller design and cost function for the trajectory optimization problem.
+In this paper, we propose a multifidelity approach to stochastic trajectory optimization with sample average approximation.
+The objective is to minimize the expected value of the objective function, which is chosen an an LQR controller for a mass-spring-damper system.
+A mass-spring-damper system is considered for ease of modelling, since the focus is on the multifidelity approach
+and not on the complexity of the system dynamics.
+The LQR controller attempts to find the optimal trajectory for tracking a reference position and velocity of the mass-spring-damper system.
+This section describes how the dynamics for the mass-spring-damper is set up, along with the controller design and objective function for the trajectory optimization problem.
+
 \subsection{Mass-Spring-Damper System}
-For this paper, a simplified trajectory optimization problem is considered for ease of modelling.
-The objective is to solve a reference tracking problem for the position and velocity of a simple mass-spring-damper system.
-The mass-spring-damper system has the following parameters:
-
-\begin{itemize}
-  \item $m = 5$ g : Mass
-  \item $k = 2$ N/m : Spring constant
-  \item $c = 0.5$ Ns/m : Damping coefficient
-    % \item Initial position: $x_0 = 0$ m
-    % \item Initial velocity: $\dot{x}_0 = 0$ m/s
-\end{itemize}
-
+The mass-spring-damper system is a second order system that can be modelled using Newton's second law of motion.
+The system in consideration has a mass $m=5$ g, spring constant $k=2$ N/m and damping coefficient $c=0.5$ Ns/m.
 Given an external force $F$ and the acceleration of the system $\ddot{x}$,
@@
 
-We have the state as $[x \dot{x}]^T$ containing the position and velocity of the mass, both of which are observed.
-The control input $u=F$ is applied to the mass, yielding the following state space model:
+We have the state $[x \dot{x}]^T$ that contains the position and velocity of the mass, both of which are observed.
+The control input $u=F$ is applied to the mass, yielding the following state-space model:
 
@@
 The continuous-time system can be converted to discrete time using such that $x_{k+1} = A x_k + B u_k$ and $y_k = C x_k$ for a chosen sampling time $T_s$.
-The $A$ and $B$ matrices for the discrete time system are then used to formulate the cost function for the LQR problem.
+The $A$ and $B$ matrices for the discrete time system are then used to formulate the objective function for the LQR problem.
 
 We set up an LQR controller to track the reference position and velocity of the mass-spring-damper system.
-An LQR controller was chosen because it has a quadratic cost function that can be minimized analytically and gives us a
+An LQR controller was chosen because it has a quadratic objective function that can be minimized analytically and gives us a
 good reference for the trajectory optimization problem.
 
-The LQR controller minimizes the cost function
+The LQR controller minimizes the objective function.
 
@@
 
-Where
-
-\begin{itemize}
-  \item $\Delta u_0, \Delta u_1 ... \Delta u_{N-1}$ : control inputs at each timestep of the trajectory
-  \item $J$ : total cost to be minimized
-  \item $N$ : time horizon
-  \item $x_k$ : state vector at time step $k$
-  \item $u_k$ : control input vector at time step $k$
-  \item $\Delta u_k = u_k - u_{k-1}$ : change in control input at time step $k$
-  \item $r_k$ : reference output at time step $k$
-  \item $e_k = y_k - r_k = Cx - r_k$ : error in the output at time step $k$
-  \item $Q_e$ : error cost matrix
-  \item $R$ : control cost matrix
-\end{itemize}
-
-Note that we can add a terminal penalty using the DARE solution for the LQR problem, but for simplicity we will not consider it here.
+Where $\Delta u_0, \Delta u_1 ... \Delta u_{N-1}$ are the control inputs at each timestep of the trajectory,
+$J$ is the objective function to be minimized, $N$ is the prediction horizon, $x_k$ is the state vector at time step $k$,
+$u_k$ is the control input vector at time step $k$, $\Delta u_k = u_k - u_{k-1}$ is the change in control input at time step $k$,
+$r_k$ is the reference output at time step $k$, $e_k = y_k - r_k = Cx - r_k$ is the error in the output at time step $k$,
+$Q_e$ is the error penalty matrix and $R$ is the control penalty matrix.
+Note that we can add a terminal penalty using the DARE solution for the LQR problem, but for simplicity, we will not consider it here.
 
@@
 
-With the resulting state space model
+With the resulting state-space model
 
@@
 For simplicity, from now on we refer to $x_k^{ext}$ as $x_k$, $A^{ext}$ as $A$ and $B^{ext}$ as $B$.
-The cost function can now be written as
+The objective function can now be written as
 
@@
     0 & 0 & \ldots & 0
-  \end{bmatrix} \text{with size } Nn_x \times Nn_x \\
+  \end{bmatrix} & \in \mathbb{R}^{Nn_x \times Nn_x} \\
   \bar{R} &=
@@
     0 & 0 & \ldots & R
-  \end{bmatrix} \text{with size } Nn_u \times Nn_u
+  \end{bmatrix} & \in \mathbb{R}^{Nn_u \times Nn_u}
 \end{aligned}
@@
 
-The cost function can then be written as a quadratic function of the control inputs $U$
+The objective function can then be written as a quadratic function of the control inputs $U$
 
@@
 
-The cost function can now be written as a quadratic function of the control inputs $U$
+The objective function can now be written as a quadratic function of the control inputs $U$
 
@@
 
-Which has a gradient with respect to $U$ as
+This has a gradient with respect to $U$ as
 
@@
 
-Assuming minimum exists, we can set the gradient to zero to get the optimal control input $U^*$
+Assuming a minimum exists, we can set the gradient to zero to obtain the optimal control input $U^*$
 
@@
 
-\subsection{Stochastic LQR Controller} \label{stochastic_lqr}
+\subsection{Stochastic LQR Controller} \label{lqr_ctrl_stoch}
 
 We consider the case where the initial state $x_0$ is stochastic, with mean $\mu$ and covariance $\Sigma$.
-Hence, the state vector $X$ is stochastic, since it is propagating the randomness of $x_0$ forward.
-The control input vector $U$ is deterministic. The cost function $J$ is now a random variable,
+Hence, the state vector $X$ is stochastic, as it propagates the randomness of $x_0$ forward.
+The control input vector $U$ is deterministic. The objective function $J$ is now a random variable,
 with an expected value $\mathbb{E}[J]$ and variance $Var[J]$.
@@
 
-The expected value of the cost function over all possible realizations of $x_0$ is
+The expected value of the objective function on all possible realizations of $x_0$ is
 
@@
 
-The expectation of cost function can be written as
+The expectation of objective function can be written as
 
@@
 
-Assuming minimum exists, we can set the gradient to zero to get the optimal control input $U^*$ that minimizes the expected value of the cost function
+Assuming a minimum exists, we can set the gradient to zero to obtain the optimal control input $U^*$ that minimizes the expected value of the objective function.
 
@@
 
-\subsubsection{Variance and Covariance of cost function in terms of initial state}
+\subsubsection{Variance and Covariance of objective function in terms of initial state}
 
@@
 
-The expected value of the cost function can be rewritten as
+The expected value of the objective function can be rewritten as
 
@@
 
-Where
+where
 
@@
 
-To determine the covariance between the costs of two solutions $U_1$ and $U_2$, in different cost functions
-$J_1$ and $J_2$, we need to first rewrite the cost functions to differentiate between them.
+To determine the covariance between the objectives of two solutions $U_1$ and $U_2$, in different objective functions
+$J_1$ and $J_2$, we need to rewrite the objective functions to differentiate between them.
 
-For $U_1$, the cost function is
+For $U_1$, the objective function is
 
@@
 
-Similarly, for $U_2$, the cost function is
+Similarly, for $U_2$, the objective function is
 
@@
 % - https://stats.stackexchange.com/questions/35084/covariance-of-a-linear-and-quadratic-form-of-a-multivariate-normal
-the covariance between their costs can be calculated as
+The covariance between their objectives can be calculated as
 
@@
 \section{Monte Carlo with Control Variates}
-TODO: explain why we need control variates
 
-This section formulates the stochastic trajectory optimization problem as a Monte Carlo problem with a low fidelity control variate.
+The objective of this paper is to reduce the variance of the objective function estimation under sample average approximation.
+A Monte Carlo(MC) simulation can be used to estimate the expected value of the objective function, but a control variate(CV) approach
+would reduce the variance with the same computational cost. In the absence of an analytical expression of the expected value of the
+objective function, an approximate control variate(ACV) approach is also proposed. The CV estimator will also serve as
+the benchmark for the ACV estimator, since it is the best possible variance reduction that can be achieved with the ACV estimator.
 
-\subsection{Comparing optimums of high and low fidelity models} \label{comp_opt}
+In order to formulate the CV and ACV approaches, we need a low-fidelity(LF) model of the system that correlates well
+with the high-fidelity(HF) model. The LF model should be computationally cheaper to evaluate than the
+HF model, and hence can be used as a control variate to reduce the variance of the objective function
+estimation.
 
-To determine if the low fidelity model can be used as a control variate, we need to determine if it correlates with the high fidelity model.
+For this paper, the HF model is the LQR controller described in section \ref{lqr_ctrl_stoch}, with a time
+discrectization of 0.05s and the LF model is the same LQR controller but with a time discrectization of 0.5s.
+The solution for the LQR controller with the LF model contains fewer elements than the HF model due to there being
+fewer time steps for the same prediction horizon. For CV/ACV construction, we need to be able to use the same
+HF control input in the LF model for the estimators to be effective.
 
-System Setup:
-\begin{itemize}
-  \item $J_h(u, x_0)$: High fidelity model/cost function
-  \item $J_l(u, x_0)$: Low fidelity model/cost function
-  \item $u_h$: Control input in high fidelity model, $n_{u_h} \times 1$
-  \item $u_l$: Control input in low fidelity model, $n_{u_l} \times 1$
-  \item $x_0$: Random initial state
-  \item $u_h^*$: High fidelity optimum
-  \item $u_l^*$: Low fidelity optimum
-\end{itemize}
+Hence, it is necessary to downsample the HF solution if we want to use it with the LF model in the CV/ACV estimators.
 
-$u_h$ and $u_l$ do not have the same dimensions, so to compare them we need to scale them to the same dimensions.
+To downsample $U_h \rightarrow U_l$, we can either restrict the dimensions of $U_h$ to match $U_l$ or average the elements of $U_h$ to match the dimensions of $U_l$.
+In case of restriction, $U_{hlr} = T_{hlr} U_h$ while in case of averaging, $U_{hla} = T_{hla} U_h$.
+$T_{hlr}$, $T_{hla} \in \mathbb{R}^{n_{U_l} \times n_{U_h}}$.
 
-To upsample $u_l \rightarrow u_h$, we repeat the elements of $u_l$ to match the dimensions of $u_h$.
-This can be seen as multiplying $u_l$ with a matrix $T_{lh}$ of size $n_{u_h} \times n_{u_l}$:
-
-\begin{equation}
-  T_{lh} =
-  \begin{bmatrix}
-    1 & 0 & \ldots & 0 \\
-    1 & 0 & \ldots & 0 \\
-    \vdots & \vdots & \ddots & \vdots \\
-    0 & 1 & \ldots & 0 \\
-    0 & 1 & \ldots & 0 \\
-    \vdots & \vdots & \ddots & \vdots \\
-    0 & 0 & \ldots & 1 \\
-    0 & 0 & \ldots & 1 \\
-    \vdots & \vdots & \ddots & \vdots \\
-  \end{bmatrix}
-\end{equation}
-
-where the number of repeating ones in each column is determined by the ratio $r_{hl} = n_{u_h} / n_{u_l}$.
-This gives us $u_{lh} = T_{lh} u_l$, which can be compared with $u_h$.
-
-To downsample $u_h \rightarrow u_l$, we can either restrict the dimensions of $u_h$ to match $u_l$ or average the elements of $u_h$ to match the dimensions of $u_l$.
-
-In case of restriction, we can multiply $u_h$ with a matrix $T_{hlr}$ of size $n_{u_l} \times n_{u_h}$:
-
-\begin{equation}
+$$
 T_{hlr} =
@@
 \end{bmatrix}
-\end{equation}
+$$
 
-Where the number of columns after which the next one is selected is determined by the ratio $r_{hl} = n_{u_h} / n_{u_l}$.
-This gives us $u_{hlr} = T_{hlr} u_h$, which can be compared with $u_l$.
+Where the number of columns after which the next one is selected is determined by the ratio $r_{hl} = n_{U_h} / n_{U_l}$.
 
-To downsample $u_h$ to $u_l$ by averaging, we need the ratio $r_{lh} = n_{u_h} / n_{u_l}$.
-We can then multiply $u_h$ with a matrix $T_{hla}$ of size $n_{u_l} \times n_{u_h}$:
-
-\begin{equation}
+$$
 T_{hla} =
@@
 \end{bmatrix}
+$$
+
+Where $r_{lh} = n_{U_l} / n_{U_h}$ is repeated $r_{hl}$ times in each row.
+
+Let $J_h$ and $J_l$ be the objective functions of the HF and LF models respectively. Then, for us to construct a
+CV/ACV estimator, we need to consider the correlation between them. With a higher correlation, we get a greater variance reduction
+We calculate the Pearson correlation coefficient $\rho$ using the covariance of the objectives
+as described in section \ref{lqr_ctrl_stoch}.
+
+\begin{equation}
+  \rho(J(U_h), J(U_l)) = \frac{\text{Cov}(J(U_h), J(U_l))}{\sigma_h \sigma_l}
 \end{equation}
 
-Where $r_{lh}$ is repeated $r_{hl}$ times in each row.
-This gives us $u_{hla} = T_{hla} u_h$, which can be compared with $u_l$.
+\subsection{CV and ACV Construction}
 
-\subsection{Correlation between high and low fidelity models}
+First, we define the Monte Carlo estimator $S_n^{MC}(J_h(U_h, x_0))$ for the HF objective as a baseline.
 
-$u_h$ and $u_l$ each have $n_{u_h}$ and $n_{u_l}$ dimensions respectively, which represent the time steps of the simulation.
-After converting them to the same dimensions as done in section \ref{comp_opt},
-the cost for some realizations of the initial state $x_0$ can be computed.
-Hence, a set of costs of both high and low fidelity control inputs can be computed for a certain model/cost function.
-The correlation coefficient $\rho$ can then be computed between the distribution of the cost function values.
-To illustrate this, assume there are $N$ realizations of the initial state $x_0$, and $J$ is the cost function.
-$u_h$ and $u_l$ are the control inputs for the high and low fidelity models respectively, where the dimensions have been
-converted to the same number as described in section \ref{comp_opt}. $\mu_h$ and $\mu_l$ are the mean costs
-of the high and low fidelity control inputs respectively, and $\sigma_h$ and $\sigma_l$ are the standard deviations of
-the costs of the high and low fidelity control inputs respectively.
+$$
+S_n^{MC}(J_h(U_h, x_0)) = \frac{1}{n} \sum_{i=1}^{n} J_h(U_h, x_0^{(i)})
+$$
 
-\begin{equation}
-  \rho(J(u_h), J(u_l)) = \frac{1}{N-1} \sum_{i=1}^{N} \left( \frac{J(u_h, x_0^i) - \mu_h}{\sigma_h} \right) \left( \frac{J(u_l, x_0^i) - \mu_l}{\sigma_l} \right)
-\end{equation}
+Where $n$ is the number of samples of $x_0$ used in the MC estimate.
 
-This can be simplified to:
+The CV estimator $S_{n_{cv}}^{CV}(J_h, J_l)$ is constructed using both the HF and LF objectives, and the
+analytical expectation of the LF objective function.
 
-\begin{equation}
-  \rho(J(u_h), J(u_l)) = \frac{\text{Cov}(J(u_h), J(u_l))}{\sigma_h \sigma_l}
-\end{equation}
+$$
+S_{n_{cv}}^{CV}(J_h, J_l) = S_{n_{cv}}^{MC}(J_h(U_h, x_0)) + \alpha (S_{n_{cv}}^{MC}(J_l(U_{hla}, x_0)) - \mathbb{E}[J_l(U_{hla}, x_0)])
+$$
 
-This correlation coefficient $\rho$ can be calculated using the MATLAB function \href{https://www.mathworks.com/help/matlab/ref/corrcoef.html#bunkaln}{\texttt{corrcoef}}.
+Choosing $n_{cv}$ samples of $x_0$ to equal the cost of $S_n^{MC}(J_h(U_h, x_0))$ and reusing the same samples for the HF and LF objectives.
+The control variate coefficient $\alpha$ can be calculated analytically using the actual mean and covariance of $x_0$
+and the equations in section \ref{lqr_ctrl_stoch}.
 
-The correlation can also be calculated analytically using the covariance of the costs in the high fidelity objective function and the low fidelity objective function.
-Let
+$$
+\alpha = -\frac{\text{Cov}(J_h(U_h, x_0), J_l(U_{hla}, x_0))}{\text{Var}[J_l(U_{hla}, x_0)]}
+$$
 
-\begin{itemize}
-  \item $J_h$: high fidelity cost function
-  \item $J_l$: low fidelity cost function
-  \item $U_1$: high fidelity control input
-  \item $U_2$: downsampled high fidelity control input $= T_{hlr} U_1$ or $T_{hla} U_1$
-  \item $J_1 = J_h(U_1, x_0)$
-  \item $J_2 = J_l(U_2, x_0)$
-  \item $\sigma_1$: standard deviation of $J_1$, $=\sqrt{\text{Var}[J_1]}$
-  \item $\sigma_2$: standard deviation of $J_2$, $=\sqrt{\text{Var}[J_2]}$
-\end{itemize}
+Finally, the ACV estimator $S_{n_{acv},m_{acv}}^{ACV}(J_h, J_l)$ is constructed using both the HF and LF objectives but with
+a statistical estimate of the expectation of the LF objective function.
 
-Then correlation coefficient between $J_1$ and $J_2$ is
+$$
+S_{n_{acv},m_{acv}}^{ACV}(J_h, J_l) = S_{n_{acv}}^{MC}(J_h(U_h, x_0)) + \alpha (S_{n_{acv}}^{MC}(J_l(U_{hla}, x_0)) - S_{m_{acv}}^{MC}(J_l(U_{hla}, x_0)))
+$$
 
-\begin{equation}
-  \begin{aligned}
-    \rho(J_1, J_2) &= \frac{\text{Cov}(J_1, J_2)}{\sigma_1 \sigma_2} \\
-    &= \frac{\text{Cov}(J_1, J_2)}{\sqrt{\text{Var}[J_1]} \sqrt{\text{Var}[J_2]}}
-  \end{aligned}
-\end{equation}
+Choosing $n_{acv} + m_{acv}$ samples of $x_0$ to equal the cost of $S_n^{MC}(J_h(U_h, x_0))$ and reusing the same samples
+for the HF and LF objectives in the first two terms. The coefficient $\alpha$ can be calculated statistically with $n_{acv}$ samples of $x_0$.
 
-Where $\text{Cov}(J_1, J_2)$, $\text{Var}[J_1]$ and $\text{Var}[J_2]$ can be calculated using the equations in section \ref{lqr_ctrl}.
+We expect the CV estimator to have a lower variance than the MC estimator. Howevever, since it is not possible
+to have an analytical expression for the expected value of most objective functions used for trajectory optimization,
+especially with constraints, the CV estimator is not always possible to construct. The ACV estimator is a compromise
+that converges to the variance of the CV estimator as $m_{acv} \rightarrow \infty$.
 
-\subsection{Correlation Experiments}
+\subsection{Numerical optimization}
 
-We conducted experiments to check the correlation in 2 scenarios:
-\begin{itemize}
-  \item Perturbation in the direction of steepest ascent
-  \item Along the path of a numerical optimizer
-\end{itemize}
+To optimize trajectory using LQR, to obtain the optimal control inputs $U_h^*$ under stochastic initial conditions $x_0$,
+we need to minimize the expected value of the objective function $\mathbb{E}[J_h(U_h, x_0)]$. We do this using off-the-shelf
+numerical optimization algorithms present in MATLAB, such as fmincon, fminunc, etc. An important point to consider
+is that the variance of the objective function estimation depends on the number of samples of $x_0$ used as well as
+the correlation between the HF and LF objectives. We exploit the fact that as the numerical optimizer iterates, it computes
+$U_h$ at every iteration and for us to achieve the maximum correlation between the HF and LF objectives, we can use
+choose a $U_h$ which when downsampled to $U_l$ gives us the highest corelation $\rho(J_h(U_h, x_0), J_l(U_{hla}, x_0))$.
+Hence, at each iteration of the CV/ACV estimators we use the $U_h$ at the current iteration while we use $U_{hla}$ from
+the iteration with the highest correlation. This $U_{hla}$ at the iteration with the highest correlation is
+$U_{hla}^{max}$. The LF objective is then calculated using this $U_{hla}^{max}$ to be used in the CV/ACV estimators to get the maximum variance reduction.
 
-\subsubsection{Perturbation in the direction of steepest ascent}
+For numerical optimization using the CV estimator, we save the $U_{hla}^{(k)}$ for every iteration $k$ in an array
+so that they can be used to calculate the correlation in all future iterations.
 
-Since the solution of the cost function is determined analytically, we cannot use it to determine the direction of steepest ascent.
-The gradient at the solution will be 0. To get the direction of steepest ascent, the solution was perturbed slightly in random directions and the direction with the highest gradient
-was chosen as the direction of steepest ascent.
+\begin{algorithm}
+  \caption{Numerical optimization with CV estimator}
+  \begin{algorithmic}
+    \Require $J_h, J_l, n_{cv}, U_0, x_0 \mu, \Sigma, k_{max}$
+    \State $k \gets 0$ (iteration counter)
+    \State $U \gets U_0$ initialize the numerical optimizer
+    \While{numerical optimizer has not converged $\wedge$ $k < k_{max}$}
+    \State $U_h \gets$ $U$ from the numerical optimizer
+    \State $U_{hla}^{(k)} \gets T_{hla} U_h$
+    \State $U_{hla}^{max} \gets \underset{U_{hla}(j)} \max \{\rho(J_h(U_h, x_0), J_l(U_{hla}^{(j)}, x_0))\}_{j=1}^k$
+    \State $\text{Cov}_{hl} \gets \text{Cov}(J_h(U_h, x_0), J_l(U_{hla}^{max}, x_0))$ calculated analytically using $\mu, \Sigma$
+    \State $\text{Var}_l \gets \text{Var}[J_l(U_{hla}^{max}, x_0)]$ calculated analytically using $\mu, \Sigma$
+    \State $\alpha \gets -\frac{\text{Cov}_{hl}}{\text{Var}_l}$
+    \State $\mathbb{E}[J_l(U_{hla}^{max}, x_0)] \gets$ calculated analytically using $\mu, \Sigma$
+    \State $S_{n_{cv}}^{CV}(J_h, J_l) \gets S_{n_{cv}}^{MC}(J_h(U_h, x_0)) + \alpha (S_{n_{cv}}^{MC}(J_l(U_{hla}^{max}, x_0)) - \mathbb{E}[J_l(U_{hla}^{max}, x_0)])$
+    \State $k \gets k + 1$
+    \State $S_{n_{cv}}^{CV}(J_h, J_l)$ is passed to the numerical optimizer for computing gradients for the next iteration
+    \EndWhile
+  \end{algorithmic}
+\end{algorithm}
 
-\textbf{Perturbation in the high fidelity model/cost function}
+Numerical optimization using ACV is similar to CV, but all the terms are calculated statistically using $n_{acv}$ and separate
+$m_{acv}$ samples of $x_0$.
 
-When perturbing in the high fidelity model/cost function, the perturbed high fidelity solution was downsampled to low fidelity
-dimensions by either restricting or averaging as described in section \ref{comp_opt}.
+\begin{algorithm}
+  \caption{Numerical optimization with ACV estimator}
+  \begin{algorithmic}
+    \Require $J_h, J_l, n_{acv}, m_{acv}, U_0, x_0, k_{max}$
+    \State $k \gets 0$ (iteration counter)
+    \State $U \gets U_0$ initialize the numerical optimizer
+    \While{numerical optimizer has not converged $\wedge$ $k < k_{max}$}
+    \State $U_h \gets$ $U$ from the numerical optimizer
+    \State $U_{hla}^{(k)} \gets T_{hla} U_h$
+    \State $U_{hla}^{max} \gets \underset{U_{hla}(j)} \max \{\rho(J_h(U_h, x_0), J_l(U_{hla}^{(j)}, x_0))\}_{j=1}^k$
+    \State $\text{Cov}_{hl} \gets \text{Cov}(J_h(U_h, x_0), J_l(U_{hla}^{max}, x_0))$ calculated statistically using $n_{acv}$ samples
+    \State $\text{Var}_l \gets \text{Var}[J_l(U_{hla}^{max}, x_0)]$ calculated statistically using $n_{acv}$ samples
+    \State $\alpha \gets -\frac{m_{acv}}{m_{acv} + n_{acv}} \frac{\text{Cov}_{hl}}{\text{Var}_l}$ calculated statistically using $n_{acv}$ samples
+    \State $\mathbb{E}[J_l(U_{hla}^{max}, x_0)] \gets \frac{1}{m_{acv}} \sum_{i=n_{acv}}^{n_{acv}+m_{acv}} J_l(U_{hla}^{max}, x_0^{(i)})$
+    \State $S_{n_{acv},m_{acv}}^{ACV}(J_h, J_l) \gets S_{n_{acv}}^{MC}(J_h(U_h, x_0)) + \alpha (S_{n_{acv}}^{MC}(J_l(U_{hla}^{max}, x_0)) - S_{m_{acv}}^{MC}(J_l(U_{hla}^{max}, x_0)))$
+    \State $k \gets k + 1$
+    \State $S_{n_{acv},m_{acv}}^{ACV}(J_h, J_l)$ is passed to the numerical optimizer for computing gradients for the next iteration
+    \EndWhile
+  \end{algorithmic}
+\end{algorithm}
 
-Hence, we have the following important variables:
-
-\begin{itemize}
-  \item $J_h$: High fidelity model/cost function
-  \item $J_l$: Low fidelity model/cost function
-  \item $u_h^*$: High fidelity optimum
-  \item $p_h$: High fidelity perturbation
-  \item $u_h = u_h^* + p_h$: Perturbed high fidelity solution
-  \item $u_{hlr} = T_{hlr} u_h$: Perturbed high fidelity solution downsampled using restriction:
-  \item $u_{hla} = T_{hla} u_h$: Perturbed high fidelity solution downsampled using averaging
-\end{itemize}
-
-Comparing $J_h(u_h, x_0)$ and $J_l(u_{hlr}, x_0)$
-
-Comparing $J_h(u_h, x_0)$ and $J_l(u_{hla}, x_0)$
-
-\textbf{Perturbation in the low fidelity model/cost function}
-
-When perturbing in the low fidelity model/cost function, the perturbed high fidelity solution was determined
-by converting the low fidelity perturbation to high fidelity dimensions as described in section \ref{comp_opt}.
-
-Hence, we have the following important variables:
-
-\begin{itemize}
-  \item $J_h$: High fidelity model/cost function
-  \item $J_l$: Low fidelity model/cost function
-  \item $u_l^*$: Low fidelity optimum
-  \item $p_l$: Low fidelity perturbation
-  \item $u_l = u_l^* + p_l$: Perturbed low fidelity solution
-  \item $u_{lh} = T_{lh} u_l$: Perturbed high fidelity solution upsampled
-\end{itemize}
-
-Comparing $J_h(u_{lh}, x_0)$ and $J_l(u_l, x_0)$
-
-\subsubsection{Along the path of a numerical optimizer}
-
-Since in a realistic scenario, one would be optimizing using a numerical optimizer, and desire to reach the high fidelity optimum,
-we conducted experiments to check the correlation along the path of a numerical optimizer in the high fidelity model/cost function.
-
-Here, we warm start with the low fidelity optimum, and check the correlation by downsampling the high fidelity solution using averaging.
-
-Hence, we have the following important variables:
-
-\begin{itemize}
-  \item $J_h$: High fidelity model/cost function
-  \item $J_l$: Low fidelity model/cost function
-  \item $u_h$: High fidelity solution
-  \item $u_{hlr} = T_{hlr} u_h$: High fidelity solution downsampled using restriction
-  \item $u_{hla} = T_{hla} u_h$: High fidelity solution downsampled using averaging
-\end{itemize}
-
-Comparing $J_h(u_h, x_0)$ and $J_h(u_{hlr}, x_0)$
-
-Comparing $J_h(u_h, x_0)$ and $J_l(u_{hla}, x_0)$
-
-It is also possible to compute the correlation between $J_h(u_h, x_0)$ and $J_l(u_{hlr}, x_0)$ for every iteration of the numerical optimizer.
-This helps us determine if there is a certain iteration of $J_l(u_{hlr}, x_0)$ that correlates well with all iterations of $J_h(u_h, x_0)$.
-The same comparison can be made with $J_l(u_{hla}, x_0)$.
-
-Comparing correlation between $J_h(u_h, x_0)$ and $J_h(u_{hlr}, x_0)$ across all iterations
-
-Comparing correlation between $J_h(u_h, x_0)$ and $J_l(u_{hla}, x_0)$ across all iterations
-
-\subsection{Control Variate Construction}
-
-The control variate is constructed using the low fidelity model/cost function, since we have found that there is some correlation between $J_h(u_h, x_0)$ and $J_l(u_{hla}, x_0)$.
-Here we are only using the averaged downsampled high fidelity solution $u_{hla}$ as the control variate.
-The important variables for constructing the control variate are:
-
-\begin{itemize}
-  \item $x_0$: Random initial state
-  \item $J_h$: High fidelity model/cost function
-  \item $J_l$: Low fidelity model/cost function
-  \item $u_h$: High fidelity solution
-  \item $u_{hla} = T_{hla} u_h$: High fidelity solution downsampled using averaging
-  \item $S_n^{MC}(J_h(u_h, x_0))$: Monte Carlo estimate of the high fidelity cost function
-  \item $S_n^{MC}(J_l(u_{hla}, x_0))$: Monte Carlo estimate of the low fidelity cost function
-  \item $\alpha$: Control variate coefficient
-\end{itemize}
-
-The control variate coefficient $\alpha$ can be calculated as
-
-\begin{equation}
-  \alpha = -\frac{\text{Cov}(J_h(u_h, x_0), J_l(u_{hla}, x_0))}{\text{Var}[J_l(u_{hla}, x_0)]}
-\end{equation}
-
-Hence, the control variate is
-
-\begin{equation}
-  S_n^{CV}(J_h, J_l) = S_n^{MC}(J_h(u_h, x_0)) + \alpha (S_n^{MC}(J_l(u_{hla}, x_0)) - \mathbb{E}[J_l(u_{hla}, x_0)])
-\end{equation}
-
-Where the samples of $x_0$ used in $J_h(u_h, x_0)$ and $J_l(u_{hla}, x_0)$ are the same.
-
-Note that $\text{Cov}(J_h(u_h, x_0), J_l(u_{hla}, x_0))$, $\text{Var}[J_l(u_{hla}, x_0)]$ and $\mathbb{E}[J_l(u_{hla}, x_0)]$ are calculated analytically using the equations in section \ref{stochastic_lqr}.
-
-\subsection{Numerical optimizer using Control Variate}
-
-The numerical optimizer was run in 3 different scenarios
-
-\begin{itemize}
-  \item $S_n^{MC}(J_h(U_h, x_0))$: MC estimate of the high fidelity cost function
-  \item $S_n^{CV}(J_h(U_h, x_0), J_l(U_{hla}, x_0))$: CV estimate with $u_{hla}$ at the current iteration
-  \item $S_n^{CV_{max}}(J_h(U_h, x_0), J_l(U_{hla}^{max}, x_0))$:CV estimate where $u_{hla}^{max}$ is at the iteration with the highest correlation.
-\end{itemize}
-
-This is how the whole routine works, step-by-step:
-
-Initialization:
-\begin{itemize}
-  \item $U_l^* \gets$ Analytical solution of low fidelity model
-  \item $U_{h_0} \gets T_{lh} U_l^*$
-\end{itemize}
-
-% For $S_n^{CV}$ the numerical optimizer is run for $i^{max}$ iterations, where in each iteration $i$, the objective is calculated:
-% \begin{equation}
-%   S_n^{CV}(J_h, J_l) = S_n^{MC}(J_h(U, x_0)) + \alpha (S_n^{MC}(J_l(U_{hla}, x_0)) - \mathbb{E}[J_l(U_{hla}, x_0)])
-% \end{equation}
-
-% Where
-% \begin{itemize}
-%   \item $n$: Number of samples of $x_0$ used in the MC estimate to equal the cost of $S^{MC}(J_h(U, x_0))$ with all samples
-%   \item $U_{hla} = T_{hla} U$
-%   \item $\alpha = -\frac{\text{Cov}(J_h(U, x_0), J_l(U_{hla}, x_0))}{\text{Var}[J_l(U_{hla}, x_0)]}$ is calculated analytically
-%   \item $\mathbb{E}[J_l(U_{hla}, x_0)]$ is calculated analytically
-% \end{itemize}
-
-% For $S_n^{CV_{max}}$ the numerical optimizer is run for $i^{max}$ iterations, where in each iteration $i$, the objective is calculated:
-% \begin{equation}
-%   S_n^{CV_{max}}(J_h, J_l) = S_n^{MC}(J_h(U, x_0)) + \alpha (S_n^{MC}(J_l(U_{hla}^{max}, x_0)) - \mathbb{E}[J_l(U_{hla}^{max}, x_0)])
-% \end{equation}
-
-% Where
-% \begin{itemize}
-%   \item $n$: Number of samples of $x_0$ used in the MC estimate to equal the cost of $S^{MC}(J_h(U, x_0))$ with all samples
-%   \item $U_{hla} = T_{hla} U$
-%   \item $U_{hla}^{max} = \underset{U_{hla}(j)}{max} \{\rho(J_h(U(i), x_0^n), J_l(U_{hla}(j), x_0^n))\}_{j=1}^i$
-%   \item $\alpha = -\frac{\text{Cov}(J_h(U, x_0), J_l(U_{hla}^{max}, x_0))}{\text{Var}[J_l(U_{hla}^{max}, x_0)]}$ is calculated analytically
-%   \item $\mathbb{E}[J_l(U_{hla}^{max}, x_0)]$ is calculated analytically
-% \end{itemize}
-
-For $S_{nm}^{ACV}$ the numerical optimizer is run for $i^{max}$ iterations, where in each iteration $i$, the objective is calculated:
-\begin{equation}
-  S_{nm}^{ACV}(J_h, J_l) = S_n^{MC}(J_h(U, x_0)) + \alpha (S_n^{MC}(J_l(U_{hla}, x_0)) - S_m^{MC}(J_l(U_{hla}, x_0)))
-\end{equation}
-
-Where
-\begin{itemize}
-  \item $m$: Total number of samples of $x_0$
-  \item $n$: Number of samples of $x_0$ used in the MC estimate to equal the cost of $S_m^{MC}(J_h(U, x_0))$ (without considering the cost of $S_m^{MC}(J_l(U_{hla}, x_0))$)
-  \item $U_{hla} = T_{hla} U$
-  \item $\alpha = -\frac{\text{Cov}(J_h(U, x_0^n), J_l(U_{hla}, x_0^n))}{\text{Var}[J_l(U_{hla}, x_0^n)]}$ is calculated statistically with $n$ samples
-\end{itemize}
-
-For $S_{nm}^{ACV_{max}}$ the numerical optimizer is run for $i^{max}$ iterations, where in each iteration $i$, the objective is calculated:
-\begin{equation}
-  S_{nm}^{ACV_{max}}(J_h, J_l) = S_n^{MC}(J_h(U, x_0)) + \alpha (S_n^{MC}(J_l(U_{hla}^{max}, x_0)) - S_m^{MC}(J_l(U_{hla}^{max}, x_0)))
-\end{equation}
-
-Where
-\begin{itemize}
-  \item $m$: Total number of samples of $x_0$
-  \item $n$: Number of samples of $x_0$ used in the MC estimate to equal the cost of $S_m^{MC}(J_h(U, x_0))$ (without considering the cost of $S_m^{MC}(J_l(U_{hla}, x_0))$)
-  \item $U_{hla} = T_{hla} U$
-  \item $U_{hla}^{max} = \underset{U_{hla}(j)}{max} \{\rho(J_h(U(i), x_0^n), J_l(U_{hla}(j), x_0^n))\}_{j=1}^i$
-  \item $\alpha = -\frac{\text{Cov}(J_h(U, x_0^n), J_l(U_{hla}^{max}, x_0^n))}{\text{Var}[J_l(U_{hla}^{max}, x_0^n)]}$ is calculated statistically with $n$ samples
-\end{itemize}
-
-For these experiments, $n$, the number of samples of $x_0$ used in the MC estimate, was compared between 10, 100, 500 and 1000.
-The comparison is done between $S_n^{MC}(J_h(u_h, x_0))$ vs $S_{\frac{n}{2}}^{CV}(J_h, J_l)$ and $S_{\frac{n}{2}}^{CV_{max}}(J_h, J_l)$ to keep the number of cost function evaluations the same with the control variate estimators having a lower computational cost due to half the sammples being used in the low fidelity cost function.
-
-To compare the variance, we ran a numerical optimizer with n random samples of $x_0$ and calculated the variance for each iteration.
-
-The variance was also calculated analytically for each control variate
 \end{document}

### 2025-02-18 Est.m
split up estimator iterations

@@ -88,7 +88,7 @@ classdef Est
 				var_l = var(cost_lf_all);
 				cov_hl = cov(cost_hf_all, cost_lf_all);
 				cov_hl = cov_hl(1, 2); % only off diagonal element
-				alpha = -cov_hl / var_l;
+				alpha = -m/(m+n) * cov_hl / var_l;
 				exp_l = mean(St.LQRCost(x0_rv_ext(:, n+1:n+m), lqrsol_lf, u_hla));
 				cost = cost_hf + alpha * (cost_lf - exp_l);
 			end

### 2025-02-18 stoch_traj.m

@@ -383,22 +383,77 @@ plot_convergence_dist(data.lqrsol{1}.Uopt, U_num_hf, U_num_cv, U_num_cv_fix, ["H
 
 %% G Convergence and variance with various optimizers and sample sizes
 num_rv_samples = [10 100 500];
-num_estimator_samples = 100;
+num_estimator_samples = 50;
 u0_num = repelem(data.lqrsol{2}.Uopt, 10, 1); % warm start
 max_iters = 30;
 options = optimoptions('fminunc', 'Display', 'iter', 'OutputFcn', @outfun, "MaxIterations", max_iters, "OptimalityTolerance", 1e-12);
+
 data.hf_cost = zeros(max_iters, num_estimator_samples, length(num_rv_samples));
+data.hf_u = zeros(length(u0_num), max_iters, num_estimator_samples, length(num_rv_samples));
+for num_samples=num_rv_samples
+  fprintf("\n*** RV Samples: %d ***\n", num_samples);
+  wait_bar = waitbar(0, "Running estimators with " + num_samples + " samples");
+  for i=1:num_estimator_samples
+    waitbar(i/num_estimator_samples, wait_bar);
+    % get the RV samples
+    x0_rv = mvnrnd(x0_mean, x0_cov, num_samples)';
+    x0_rv_ext = [x0_rv; repmat(u0, 1, num_samples); repmat(ref, 1, num_samples)];
+    
+    % MC with HF
+    num_opt_iters = 0;
+    num_opt_data = zeros(size(Uopt_hf,1), max_iters);
+    num_opt_fvals = zeros(1, max_iters);
+    x0_rv_ext_mean = mean(x0_rv_ext, 2);
+    fun = @(u) mean(St.LQRCost(x0_rv_ext, data.lqrsol{1}, u));
+    Uopt_num = fminunc(fun, u0_num, options);
+    data.hf_cost(:, i, num_rv_samples == num_samples) = num_opt_fvals(1:max_iters);
+    data.hf_u(:, :, i, num_rv_samples == num_samples) = num_opt_data(:, 1:max_iters);
+  end
+  close(wait_bar);
+end
+
+%%
 data.cv_cost = zeros(max_iters, num_estimator_samples, length(num_rv_samples));
 data.cv_fix_cost = zeros(max_iters, num_estimator_samples, length(num_rv_samples));
[…]

### 2025-02-24 stoch_traj.m
fix m and n in acv being int causing issues

@@ -459,10 +459,10 @@ for num_samples=num_rv_samples
   wait_bar = waitbar(0, "Running estimators with " + num_samples + " samples");
   lf_hf_cost_ratio = 0.2; % TODO get this experimentally for each num_samples
   % calculate samples given cost ratio and HF=LF
-  n_cv = int32(num_samples / (1 + lf_hf_cost_ratio));
-  n_acv = int32(num_samples / (1 + lf_hf_cost_ratio + acv_exp_sample_factor*lf_hf_cost_ratio));
-  m_acv = int32(acv_exp_sample_factor * n_acv);
-  n_total = int32(max(n_cv+n_acv+m_acv, num_samples)); % account for weird splits like 0.999
+  n_cv = round(num_samples / (1 + lf_hf_cost_ratio));
+  n_acv = round(num_samples / (1 + lf_hf_cost_ratio + acv_exp_sample_factor*lf_hf_cost_ratio));
+  m_acv = round(acv_exp_sample_factor * n_acv);
+  n_total = round(max(n_cv+n_acv+m_acv, num_samples)); % account for weird splits like 0.999
   num_rv_samples_actual(num_rv_samples == num_samples, :) = [num_samples, n_cv, n_acv, m_acv];
   for i=1:num_estimator_samples
     waitbar(i/num_estimator_samples, wait_bar);

### 2025-02-25 artefacts/paper.tex
uce unnecessary sections

@@
 The system in consideration has a mass $m=5$ g, spring constant $k=2$ N/m and damping coefficient $c=0.5$ Ns/m.
-Given an external force $F$ and the acceleration of the system $\ddot{x}$,
-the dynamics of the system can be written as the equality of forces acting on the mass:
-
-$$
-\begin{aligned}
-  m\ddot{x} &= F - c\dot{x} - kx \\
-  \implies \ddot{x} &= \frac{F}{m} - \frac{c}{m}\dot{x} - \frac{k}{m}x
-\end{aligned}
-$$
-
-We have the state $[x \dot{x}]^T$ that contains the position and velocity of the mass, both of which are observed.
+Given an external force $F$ and the acceleration of the system $\ddot{x}$, we have the state $[x \dot{x}]^T$
+that contains the position and velocity of the mass, both of which are observed.
 The control input $u=F$ is applied to the mass, yielding the following state-space model:
@@
 
-Thus, we have a standard continuous-time LTI system $\dot{x} = A_cx + B_cu$ and $y = C_cx$, where
-
-\begin{equation}
-  \begin{aligned}
-    A_c &=
-    \begin{bmatrix}
-      0 & 1 \\
-      -\frac{k}{m} & -\frac{c}{m}
-    \end{bmatrix} \\
-    B_c &=
-    \begin{bmatrix}
-      0 \\
-      \frac{1}{m}
-    \end{bmatrix} \\
-    C_c &=
-    \begin{bmatrix}
-      1 & 0 \\
-      0 & 1
-    \end{bmatrix}
-  \end{aligned}
-\end{equation}
+Thus, we have a standard continuous-time LTI system $\dot{x} = A_cx + B_cu$ and $y = C_cx$, which can be discretized
+into a discrete-time LTI system $x_{k+1} = Ax_k + Bu_k$ and $y_k = Cx_k$ using a sampling time $T_s$. We can now use
+this discrete-time system to set up the LQR controller.
 
@@
 
-With the resulting state-space model
-
-\begin{equation}
-  \begin{aligned}
-    x_{k+1}^{ext} &=
-    \begin{bmatrix}
-      A & B & 0 \\
-      0 & \mathbb{I}_{n_u \times n_u} & 0 \\
-      0 & 0 & \mathbb{I}_{n_r \times n_r}
-    \end{bmatrix}
-    \begin{bmatrix}
-      x_k \\
-      u_{k-1} \\
-      r_k
-    \end{bmatrix} +
-    \begin{bmatrix}
-      B \\
-      \mathbb{I}_{n_u \times n_u} \\
-      0
-    \end{bmatrix} \Delta u_k \\
-  \end{aligned}
-\end{equation}
-
-with
+And stack the state vectors $x_k^{ext}$ for all time steps $k$ to get the extended state vector $X$.
+The control input is also stacked to get the control input vector $U$.
 
 $$
-\begin{aligned}
-  A^{ext} &=
+X =
 \begin{bmatrix}
-    A & B & 0 \\
-    0 & \mathbb{I}_{n_u \times n_u} & 0 \\
-    0 & 0 & \mathbb{I}_{n_r \times n_r}
-  \end{bmatrix} \\
-  B^{ext} &=
+  x_1^{ext} \\
+  x_2^{ext} \\
+  \vdots \\
+  x_N^{ext}
+\end{bmatrix}  \in \mathbb{R}^{Nn_x \times 1}, \quad
+U =
 \begin{bmatrix}
-    B \\
-    \mathbb{I}_{n_u \times n_u} \\
-    0
-  \end{bmatrix}
-\end{aligned}
+  \Delta u_0 \\
+  \Delta u_1 \\
+  \vdots \\
+  \Delta u_{N-1}
+\end{bmatrix}  \in \mathbb{R}^{Nn_u \times 1}
 $$
@@
 Similary, $Qe$ with the extended state $x_k^{ext}$ transforms to $Q$ such that $e_k^T Q_e e_k = x_k^{ext^T} Q x_k^{ext}$.
-
-$$
-\begin{aligned}
-  e_k^T Q_e e_k &= (y_k - r_k)^T Q_e (y_k - r_k) \\
-  &= \left(
-    \begin{bmatrix} C & 0 & -\mathbb{I}_{n_r \times n_r}
-    \end{bmatrix}
-    \begin{bmatrix}
-      x_k \\
-      u_{k-1} \\
-      r_k
-  \end{bmatrix}\right)^T Q_e
-  \left(
-    \begin{bmatrix} C & 0 & -\mathbb{I}_{n_r \times n_r}
-    \end{bmatrix}
-    \begin{bmatrix}
-      x_k \\
-      u_{k-1} \\
-      r_k
-  \end{bmatrix}\right) \\
-  &= x_k^{ext^T}
-  \begin{bmatrix} C & 0 & -\mathbb{I}_{n_r \times n_r}
-  \end{bmatrix}^T Q_e
-  \begin{bmatrix} C & 0 & -\mathbb{I}_{n_r \times n_r}
-  \end{bmatrix} x_k^{ext} \\
-\end{aligned}
-$$
-
-\begin{equation}
-  Q =
-  \begin{bmatrix} C & 0 & -\mathbb{I}_{n_r \times n_r}
-  \end{bmatrix}^T Q_e
-  \begin{bmatrix} C & 0 & -\mathbb{I}_{n_r \times n_r}
-  \end{bmatrix}
-\end{equation}
 
@@
 
-Using the state transition formula
-
-$$
-\begin{aligned}
-  x_1 &= A x_0 + B u_0 \\
-  x_2 &= A x_1 + B u_1 = A(A x_0 + B u_0) + B u_1 = A^2 x_0 + A B u_0 + B u_1 \\
-  \vdots \\
-  x_N &= A x_{N-1} + B u_{N-1} = A^{N-1} x_0 + A^{N-2} B u_0 + A^{N-3} B u_1 + \ldots + B u_{N-1} \\
-  \implies x_k &= A^k x_0 + \sum_{i=0}^{k-1} A^{k-1-i} B u_i
-\end{aligned}
-$$
-
-We construct the matrices $X$, $U$, $S$ and $M$ such that $X = SU + Mx_0$.
-
-$$
-\begin{aligned}
-  X &=
-  \begin{bmatrix}
-    x_1^{ext} \\
-    x_2^{ext} \\
-    \vdots \\
-    x_N^{ext}
-  \end{bmatrix} & \in \mathbb{R}^{Nn_x \times 1} \\
-  U &=
-  \begin{bmatrix}
-    \Delta u_0 \\
-    \Delta u_1 \\
-    \vdots \\
-    \Delta u_{N-1}
-  \end{bmatrix} & \in \mathbb{R}^{Nn_u \times 1} \\
-  S &=
-  \begin{bmatrix}
-    B & 0 & \ldots & 0 \\
-    AB & B & \ldots & 0 \\
-    \vdots & \vdots & \ddots & \vdots \\
-    A^{N-1} B & A^{N-2} B & \ldots & B
-  \end{bmatrix} & \in \mathbb{R}^{Nn_x \times Nn_u} \\
-  M &=
-  \begin{bmatrix}
-    A \\
-    A^2 \\
-    \vdots \\
-    A^N
-  \end{bmatrix} & \in \mathbb{R}^{Nn_x \times n_x}
-\end{aligned}
-$$
-
-Similarly, the matrices $\bar{Q}$ and $\bar{R}$ are constructed
-
-$$
-\begin{aligned}
-  \bar{Q} &=
-  \begin{bmatrix}
-    Q & 0 & \ldots & 0 \\
-    0 & Q & \ldots & 0 \\
-    \vdots & \vdots & \ddots & \vdots \\
-    0 & 0 & \ldots & 0
-  \end{bmatrix} & \in \mathbb{R}^{Nn_x \times Nn_x} \\
-  \bar{R} &=
-  \begin{bmatrix}
-    R & 0 & \ldots & 0 \\
-    0 & R & \ldots & 0 \\
-    \vdots & \vdots & \ddots & \vdots \\
-    0 & 0 & \ldots & R
-  \end{bmatrix} & \in \mathbb{R}^{Nn_u \times Nn_u}
-\end{aligned}
-$$
-
-The objective function can then be written as a quadratic function of the control inputs $U$
+Using the state transition formula, we can write the stacked state vector $X$ as $X = SU + Mx_0$.
+Here $S \in \mathbb{R}^{Nn_x \times Nn_u}$ and $M \in \mathbb{R}^{Nn_x \times n_x}$ are matrices that depend on the
+state transition matrices $A$ and $B$.
+Similarly, the matrices $\bar{Q} \in \mathbb{R}^{Nn_x \times Nn_x}$ and $\bar{R} \in \mathbb{R}^{Nn_u \times Nn_u}$
+are constructed using the penalty matrices $Q$ and $R$. The objective function then can then be written as a quadratic function of the control inputs $U$
 
@@
     \mathbb{E}[J] &= \mathbb{E}[\sum_{k=0}^{N-1} x_k^T Q x_k + \Delta u_k^T R \Delta u_k] \\
-    &= \mathbb{E}[X^T \bar{Q} X + U^T \bar{R} U + x_0^T Q x_0] \\
-    &= \mathbb{E}[X^T \bar{Q} X] + \mathbb{E}[U^T \bar{R} U] + \mathbb{E}[x_0^T Q x_0] \\
-    &= \mathbb{E}[(S U + M x_0)^T \bar{Q} (S U + M x_0)] + U^T \bar{R} U + \mathbb{E}[x_0^T Q x_0] \\
-    &= \mathbb{E}[U^T S^T \bar{Q} S U + U^T S^T \bar{Q} M x_0 + x_0^T M^T \bar{Q} S U + x_0^T M^T \bar{Q} M x_0] + U^T \bar{R} U + \mathbb{E}[x_0^T] Q \mathbb{E}[x_0] + \text{tr}(Q \text{Cov}[x_0]) \\
-    &= U^T S^T \bar{Q} S U + \mathbb{E}[U^T S^T \bar{Q} M x_0] + \mathbb{E}[x_0^T M^T \bar{Q} S U] + \mathbb{E}[x_0^T M^T \bar{Q} M x_0] + U^T \bar{R} U + \mathbb{E}[x_0^T] Q \mathbb{E}[x_0] + \text{tr}(Q \text{Cov}[x_0]) \\
-    &= U^T (S^T \bar{Q} S + \bar{R}) U + U^T S^T \bar{Q} M \mathbb{E}[x_0] + \mathbb{E}[x_0^T] M^T \bar{Q} S U + \mathbb{E}[x_0^T M^T \bar{Q} M x_0] + \mathbb{E}[x_0^T] Q \mathbb{E}[x_0] + \text{tr}(Q \text{Cov}[x_0]) \\
-    &= U^T (S^T \bar{Q} S + \bar{R}) U + \mathbb{E}[x_0^T] M^T \bar{Q} S U + \mathbb{E}[x_0^T] M^T \bar{Q} S U + \mathbb{E}[x_0^T] M^T \bar{Q} M \mathbb{E}[x_0] + \text{tr}(M^T \bar{Q} M \text{Cov}[x_0]) \dots \\
-    & \quad + \mathbb{E}[x_0^T] Q \mathbb{E}[x_0] + \text{tr}(Q \text{Cov}[x_0]) \\
     &= U^T (S^T \bar{Q} S + \bar{R}) U + 2 \mathbb{E}[x_0^T] M^T \bar{Q} S U + \mathbb{E}[x_0^T] (M^T \bar{Q} M + Q) \mathbb{E}[x_0] + \text{tr}((M^T \bar{Q} M + Q) \text{Cov}[x_0]) \\
@@
 
+The solution to the stochastic LQR controller, $U^*$ can be written as a function of the initial state $x_0$.
+
+$$
+\begin{aligned}
+  U^* &= -H^{-1} q \\
+  \mathbb{E}[U^*] &= \mathbb{E}[-H^{-1} q] \\
+  &= -H^{-1} \mathbb{E}[q] \text{ ($H$ is deterministic)} \\
+  &= -H^{-1} \mathbb{E}[S^T \bar{Q} M x_0] \\
+  &= -H^{-1} S^T \bar{Q} M \mathbb{E}[x_0] \\
+  \text{Cov}[U^*] &= \text{Cov}[-H^{-1} q] \\
+  &= H^{-1} \text{Cov}[q] H^{-1} \text{ ($H$ is symmetric)} \\
+  &= H^{-1} \text{Cov}[S^T \bar{Q} M x_0] H^{-1} \\
+  &= H^{-1} S^T \bar{Q} M \text{Cov}[x_0] M^T \bar{Q} S H^{-1} \text{($\bar{Q}$ is symmetric)} \\
+\end{aligned}
+$$
+
+Thus, we aim to reduce the variance of the objective function estimate thereby reducing the covariance of the
+solution $U^*$ given a finite cost of computation.
+
 In order to formulate the CV and ACV approaches, we need a low-fidelity(LF) model of the system that correlates well
 with the high-fidelity(HF) model. The LF model should be computationally cheaper to evaluate than the
-HF model, and hence can be used as a control variate to reduce the variance of the objective function
+HF model, and hence can be evaluated with a lot more samples of $x_0$ than the HF model for the same cost.
+Thus, we use the LF model as a control variate to reduce the variance of the objective function
 estimation.
@@
     \State $\alpha \gets -\frac{m_{acv}}{m_{acv} + n_{acv}} \frac{\text{Cov}_{hl}}{\text{Var}_l}$ calculated statistically using $n_{acv}$ samples
-    \State $\mathbb{E}[J_l(U_{hla}^{max}, x_0)] \gets \frac{1}{m_{acv}} \sum_{i=n_{acv}}^{n_{acv}+m_{acv}} J_l(U_{hla}^{max}, x_0^{(i)})$
+    \State $S_{m_{acv}}^{MC}(J_l(U_{hla}^{max}, x_0)) \gets \frac{1}{m_{acv}} \sum_{i=n_{acv}}^{n_{acv}+m_{acv}} J_l(U_{hla}^{max}, x_0^{(i)})$
     \State $S_{n_{acv},m_{acv}}^{ACV}(J_h, J_l) \gets S_{n_{acv}}^{MC}(J_h(U_h, x_0)) + \alpha (S_{n_{acv}}^{MC}(J_l(U_{hla}^{max}, x_0)) - S_{m_{acv}}^{MC}(J_l(U_{hla}^{max}, x_0)))$

### 2025-03-08 Est.m
cleanup redundant fns

@@ -2,12 +2,39 @@
 
 classdef Est
 	methods(Static)
-		function cost = CvEstA(x0_rv_ext, x0_mean, x0_cov, lqrsol_hf, lqrsol_lf, n, u)
-			% MC estimator for hf
-			cost_hf = mean(St.LQRCost(x0_rv_ext(:, 1:n), lqrsol_hf, u));
+		function cost = CvEst(x0_rv_ext, x0_mean, x0_cov, lqrsol_hf, lqrsol_lf, n, u, use_best_U_lf, reset)
+			persistent Us costs_lf idx;
+			if isempty(Us)
+				Us = [];
+				costs_lf = [];
+				idx = 1;
+			end
+			if reset
+				Us = [];
+				costs_lf = [];
+				idx = 1;
+				return;
+			end
+			
+			cost_hf_all = St.LQRCost(x0_rv_ext(:, 1:n), lqrsol_hf, u);
 			% reuse same samples for LF
 			u_hla = Est.DownsampleAvg(u, 10);
-			cost_lf = mean(St.LQRCost(x0_rv_ext(:, 1:n), lqrsol_lf, u_hla));
+			cost_lf_all = St.LQRCost(x0_rv_ext(:, 1:n), lqrsol_lf, u_hla);
+			if use_best_U_lf
+				Us(:, idx) = u; % temporary init so it can be used here
+				costs_lf(:, idx) = cost_lf_all;
+				
+				% TODO: Us should be in rows to prevent transpose
+				corrs = St.CorrMulti2D(cost_hf_all', costs_lf(:, 1:idx)');
+				[~, best_idx] = max(corrs);
+				
+				u_hla = Est.DownsampleAvg(Us(:, best_idx), 10);
+				cost_lf_all = costs_lf(:, best_idx);
+                idx = idx + 1;
+			end
+			cost_hf = mean(cost_hf_all);
+			cost_lf = mean(cost_lf_all);
+			
 			% calculate optimal alpha using actual mean, cov
 			var_l = St.LQRVar(x0_mean, x0_cov, lqrsol_lf, u_hla);
 			cov_hl = St.LQRCov(x0_mean, x0_cov, lqrsol_hf, lqrsol_lf, u, u_hla);

### 2025-03-08 stoch_traj.m

@@ -227,8 +227,8 @@ costs_str = ["$J_h(u_h)$", "$J_l(u_{hlr})$"];
 plot_multifid_costs(perturb_range, mean(cost_hf,2), mean(cost_lf,2), corr_st, corr_an, title_str, costs_str, "Perturbation");
 
 %% E.1.1.2 averaging
-Uopt_lf = downsample_avg(Uopt_hf, 10); % get HF in LF by avg every 10 values of HF
-% Uopt_lf = downsample_avg(Uopt_hf, 10, true); % uncomment to visualize
+Uopt_lf = Est.DownsampleAvg(Uopt_hf, 10); % get HF in LF by avg every 10 values of HF
+% Uopt_lf = Est.DownsampleAvg(Uopt_hf, 10, true); % uncomment to visualize
 % vis_sols(Uopt_hf, Uopt_lf, data.lqrsol{1}.times, "HF obj", perturb_range, "Perturbation");
 % calculate costs and correlation for HF/LF sols for each perturbation
 cost_hf = St.LQRCostMulti(x0_rv_ext, data.lqrsol{1}, Uopt_hf);
@@ -291,8 +291,8 @@ plot_multifid_costs(1:num_opt_iters, mean(cost_hf,2), mean(cost_lf,2), corr_st, 
 plot_corr_2d(corr_2d, "Correlation between costs in $J_h(u_h)$ and $J_l(u_{hlr})$", costs_str, "num_iters_res");
 
 %% E.2.1.2 correlation with averaging
-U_lf = downsample_avg(U_hf, 10); % get LF by avg every 10 values of HF
-% U_lf = downsample_avg(U_hf, 10, true); % uncomment to visualize
+U_lf = Est.DownsampleAvg(U_hf, 10); % get LF by avg every 10 values of HF
+% U_lf = Est.DownsampleAvg(U_hf, 10, true); % uncomment to visualize
 % vis_sols(U_hf, U_lf, data.lqrsol{1}.times, "Solutions along optimizer path", 1:num_opt_iters, "Iteration");
 cost_hf = St.LQRCostMulti(x0_rv_ext, data.lqrsol{1}, U_hf);
 cost_lf = St.LQRCostMulti(x0_rv_ext, data.lqrsol{2}, U_lf);
[…]

### 2025-03-10 Cv.m
refactor cv into a class

@@ -1,0 +1,80 @@
+classdef Cv < handle
+	properties
+		Us = [];
+		costs_lf = [];
+		idx = 1;
+		x0_mean = [];
+		x0_cov = [];
+		lqrsol_hf = {};
+		lqrsol_lf = {};
+	end
+	
+	methods
+		function obj = Cv(x0_mean, x0_cov, lqrsol_hf, lqrsol_lf)
+			obj.x0_mean = x0_mean;
+			obj.x0_cov = x0_cov;
+			obj.lqrsol_hf = lqrsol_hf;
+			obj.lqrsol_lf = lqrsol_lf;
+		end
+		
+		function cost = est(obj, x0_rv_ext, n, u, use_best_U_lf)
+			% ONLY set use_best_U_lf if insider optimizer and idx is set outside
+			cost_hf_all = St.LQRCost(x0_rv_ext(:, 1:n), obj.lqrsol_hf, u);
+			u_hla = Est.DownsampleAvg(u, 10);
+			cost_lf_all = St.LQRCost(x0_rv_ext(:, 1:n), obj.lqrsol_lf, u_hla);
+			if use_best_U_lf
+				obj.Us(:, obj.idx) = u;
+				obj.costs_lf(:, obj.idx) = cost_lf_all;
+				% TODO: Us should be in rows to prevent transpose
+				corrs = St.CorrMulti2D(cost_hf_all', obj.costs_lf(:, 1:obj.idx)');
+				[~, best_idx] = max(corrs);
+				u_hla = Est.DownsampleAvg(obj.Us(:, best_idx), 10);
+				cost_lf_all = obj.costs_lf(:, best_idx);
+			end
+			cost_hf = mean(cost_hf_all);
+			cost_lf = mean(cost_lf_all);
+			var_l = St.LQRVar(obj.x0_mean, obj.x0_cov, obj.lqrsol_lf, u_hla); % analytical
+			cov_hl = St.LQRCov(obj.x0_mean, obj.x0_cov, obj.lqrsol_hf, obj.lqrsol_lf, u, u_hla); % analytical
+			alpha = -cov_hl / var_l;
+			exp_l = St.LQRExp(obj.x0_mean, obj.x0_cov, obj.lqrsol_lf, u_hla);
+			cost = cost_hf + alpha * (cost_lf - exp_l);
+		end
+		
[…]

### 2025-03-10 stoch_traj.m

@@ -308,15 +308,10 @@ plot_multifid_costs(1:num_opt_iters, mean(cost_hf,2), mean(cost_lf,2), corr_st, 
 plot_corr_2d(corr_2d, "Correlation between costs in $J_h(u_h)$ and $J_l(u_{hla})$", costs_str, "num_iters_avg");
 
 %% F num opt with CV estimator
+cv = Cv(x0_ext_mean, x0_ext_cov, data.lqrsol{1}, data.lqrsol{2});
 %% F.1 normal CV Estimator
 cv_samples = rv_samples/2;
-num_opt_iters = 0;
-num_opt_data = zeros(size(Uopt_hf,1), 100);
-Est.CvEst([], [], [], [], [], [], [], false, true); % reset the persistent vars
-fun = @(u) Est.CvEst(x0_rv_ext, x0_ext_mean, x0_ext_cov, data.lqrsol{1}, data.lqrsol{2}, cv_samples, u, false, false);
-options = optimoptions('fminunc', 'Display', 'iter', 'OutputFcn', @outfun); % finite diff
-Uopt_num = fminunc(fun, u0_num, options);
-U_hf = num_opt_data(:, 1:num_opt_iters);
+[~, U_hf] = cv.opt(u0_num, -1, -1, x0_rv_ext, cv_samples, false);
 U_num_cv = U_hf; % save for plotting later
 U_lf = Est.DownsampleAvg(U_hf, 10);
 % U_lf = Est.DownsampleAvg(U_hf, 10, true); % uncomment to visualize
@@ -332,7 +327,7 @@ corr_2d = St.LQRCorrMulti2D(x0_ext_mean, x0_ext_cov, data.lqrsol{1}, data.lqrsol
 % plot correlation
 title_str = ["$J_h$ and $J_l$", "$S_{" + cv_samples + "}^{CV}$"];
 costs_str = ["$J_h(u_h)$", "$J_l(u_{hla})$"];
-plot_multifid_costs(1:num_opt_iters, mean(cost_hf,2), mean(cost_lf,2), corr_st, corr_an, title_str, costs_str, "Iteration");
+plot_multifid_costs(1:cv.idx, mean(cost_hf,2), mean(cost_lf,2), corr_st, corr_an, title_str, costs_str, "Iteration");
[…]

### 2025-03-10 Cv.m
ouput hlas to accomodate plotting

@@ -1,6 +1,7 @@
 classdef Cv < handle
 	properties
 		Us = [];
+		U_hlas = [];
 		costs_lf = [];
 		idx = 1;
 		x0_mean = [];
@@ -31,6 +32,7 @@ classdef Cv < handle
 				u_hla = Est.DownsampleAvg(obj.Us(:, best_idx), 10);
 				cost_lf_all = obj.costs_lf(:, best_idx);
 			end
+			obj.U_hlas(:, obj.idx) = u_hla; % save bc used for plotting
 			cost_hf = mean(cost_hf_all);
 			cost_lf = mean(cost_lf_all);
 			var_l = St.LQRVar(obj.x0_mean, obj.x0_cov, obj.lqrsol_lf, u_hla); % analytical
@@ -40,9 +42,10 @@ classdef Cv < handle
 			cost = cost_hf + alpha * (cost_lf - exp_l);
 		end
 		
-		function [costs, Us] = opt(obj, u0, max_iters, tol, x0_rv_ext, n, use_best_U_lf)
+		function [costs, Us, U_hlas] = opt(obj, u0, max_iters, tol, x0_rv_ext, n, use_best_U_lf)
 			costs = zeros(max_iters, 1);
 			obj.Us = zeros(size(u0, 1), max_iters);
+			obj.U_hlas = zeros(size(u0, 1)/10, max_iters);
 			obj.costs_lf = zeros(n, max_iters); % we use only n samples out of all in x0_rv
 			obj.idx = 1;
 			
@@ -62,6 +65,7 @@ classdef Cv < handle
 			end
 			costs = costs(1:iters);
 			obj.Us = obj.Us(:, 1:iters);
+			obj.U_hlas = obj.U_hlas(:, 1:iters);
 			
 			function stop = OutFn(x, optimValues, state)
 				stop = false;
@@ -73,6 +77,7 @@ classdef Cv < handle
 				end
 			end
 			Us = obj.Us;
+			U_hlas = obj.U_hlas;
 		end
 		
 	end

### 2025-03-10 stoch_traj.m

@@ -311,10 +311,10 @@ plot_corr_2d(corr_2d, "Correlation between costs in $J_h(u_h)$ and $J_l(u_{hla})
 cv = Cv(x0_ext_mean, x0_ext_cov, data.lqrsol{1}, data.lqrsol{2});
 %% F.1 normal CV Estimator
 cv_samples = rv_samples/2;
-[~, U_hf] = cv.opt(u0_num, -1, -1, x0_rv_ext, cv_samples, false);
+[~, U_hf, U_lf] = cv.opt(u0_num, -1, -1, x0_rv_ext, cv_samples, false);
 U_num_cv = U_hf; % save for plotting later
-U_lf = Est.DownsampleAvg(U_hf, 10);
-% U_lf = Est.DownsampleAvg(U_hf, 10, true); % uncomment to visualize
+
+% U_lf = repelem(U_hla, 10, cv.idx); % uncomment to visualize
 % vis_sols(U_hf, U_lf, data.lqrsol{1}.times, "Solutions along optimizer path", 1:num_opt_iters, "Iteration");
 
 cost_hf = St.LQRCostMulti(x0_rv_ext, data.lqrsol{1}, U_hf);
@@ -331,20 +331,12 @@ plot_multifid_costs(1:cv.idx, mean(cost_hf,2), mean(cost_lf,2), corr_st, corr_an
 plot_corr_2d(corr_2d, "$S_{" + cv_samples + "}^{CV}$ opt Correlation between costs in $J_h(u_h)$ and $J_l(u_{hla})$", costs_str, "num_iters_cv");
 
 %% F.2 CV Estimator with fixed LF solution
-[~, it_max_cor] = max(sum(corr_2d, 1));
-u_lf = U_lf(:, it_max_cor);
-num_opt_iters = 0;
-num_opt_data = zeros(size(Uopt_hf,1), 100);
-Est.CvEst([], [], [], [], [], [], [], false, true); % reset the persistent vars
-fun = @(u) Est.CvEst(x0_rv_ext, x0_ext_mean, x0_ext_cov, data.lqrsol{1}, data.lqrsol{2}, cv_samples, u, true, false);
-options = optimoptions('fminunc', 'Display', 'iter', 'OutputFcn', @outfun); % finite dif
-Uopt_num = fminunc(fun, u0_num, options);
-U_hf = num_opt_data(:, 1:num_opt_iters);
[…]

### 2025-03-10 Cv.m
use cv clas for opt everywhere

@@ -80,6 +80,14 @@ classdef Cv < handle
 			U_hlas = obj.U_hlas;
 		end
 		
+		function var = var(n, u)
+			var_h = St.LQRVar(obj.x0_mean, obj.x0_cov, obj.lqrsol_hf, u);
+			u_lf = Est.DownsampleAvg(u, 10);
+			corr_hl = St.LQRCorr(obj.x0_mean, obj.x0_cov, obj.lqrsol_hf, obj.lqrsol_lf, u, u_lf);
+			var = var_h/n * (1 - corr_hl^2);
+		end
+		
+		
 	end
 	
 end

### 2025-03-10 stoch_traj.m

@@ -504,12 +504,14 @@ grid on;
 
 
 %% G Convergence and variance with various optimizers and sample sizes
-num_rv_samples = [10 100 500];
+num_rv_samples = [10];
 num_estimator_samples = 50;
 u0_num = repelem(data.lqrsol{2}.Uopt, 10, 1); % warm start
 max_iters = 30;
+tol = 1e-12;
 options = optimoptions('fminunc', 'Display', 'iter', 'OutputFcn', @outfun, "MaxIterations", max_iters, "OptimalityTolerance", 1e-12);
 
+%%
 data.hf_cost = zeros(max_iters, num_estimator_samples, length(num_rv_samples));
 data.hf_u = zeros(length(u0_num), max_iters, num_estimator_samples, length(num_rv_samples));
 for num_samples=num_rv_samples
@@ -554,7 +556,8 @@ for num_samples=num_rv_samples
     x0_rv_ext = [x0_rv; repmat(u0, 1, n_total); repmat(ref, 1, n_total)];
     
     % CV with fixed LF solution
-    [costs, Us] = Est.CvOpt(u0_num, max_iters, x0_rv_ext, x0_ext_mean, x0_ext_cov, data.lqrsol{1}, data.lqrsol{2}, n_cv, true);
+    % [costs, Us] = Est.CvOpt(u0_num, max_iters, x0_rv_ext, x0_ext_mean, x0_ext_cov, data.lqrsol{1}, data.lqrsol{2}, n_cv, true);
+    [costs, Us, ~] = cv.opt(u0_num, max_iters, tol, x0_rv_ext, n_cv, true);
     data.cv_fix_cost(:, i, num_rv_samples == num_samples) = costs;
     data.cv_fix_u(:, :, i, num_rv_samples == num_samples) = Us;
     

### 2025-03-10 Acv.m
make acv into a class

@@ -1,4 +1,4 @@
-classdef Cv < handle
+classdef Acv < handle
 	properties
 		Us = [];
 		U_hlas = [];
@@ -11,14 +11,14 @@ classdef Cv < handle
 	end
 	
 	methods
-		function obj = Cv(x0_mean, x0_cov, lqrsol_hf, lqrsol_lf)
+		function obj = Acv(x0_mean, x0_cov, lqrsol_hf, lqrsol_lf)
 			obj.x0_mean = x0_mean;
 			obj.x0_cov = x0_cov;
 			obj.lqrsol_hf = lqrsol_hf;
 			obj.lqrsol_lf = lqrsol_lf;
 		end
 		
-		function cost = est(obj, x0_rv_ext, n, u, use_best_U_lf)
+		function cost = est(obj, x0_rv_ext, n, m, u, use_best_U_lf, a_type, e_type)
 			% ONLY set use_best_U_lf if insider optimizer and idx is set outside
 			cost_hf_all = St.LQRCost(x0_rv_ext(:, 1:n), obj.lqrsol_hf, u);
 			u_hla = Est.DownsampleAvg(u, 10);
@@ -35,14 +35,28 @@ classdef Cv < handle
 			obj.U_hlas(:, obj.idx) = u_hla; % save bc used for plotting
 			cost_hf = mean(cost_hf_all);
 			cost_lf = mean(cost_lf_all);
-			var_l = St.LQRVar(obj.x0_mean, obj.x0_cov, obj.lqrsol_lf, u_hla); % analytical
-			cov_hl = St.LQRCov(obj.x0_mean, obj.x0_cov, obj.lqrsol_hf, obj.lqrsol_lf, u, u_hla); % analytical
-			exp_l = St.LQRExp(obj.x0_mean, obj.x0_cov, obj.lqrsol_lf, u_hla);
+			if strcmp(a_type, 'anly')
+				var_l = St.LQRVar(obj.x0_mean, obj.x0_cov, obj.lqrsol_lf, u_hla); % analytical
+				cov_hl = St.LQRCov(obj.x0_mean, obj.x0_cov, obj.lqrsol_hf, obj.lqrsol_lf, u, u_hla); % analytical
+			elseif strcmp(a_type, 'stat')
+				var_l = var(cost_lf_all);
+				cov_hl = cov(cost_hf_all, cost_lf_all);
+				cov_hl = cov_hl(1, 2); % only off diagonal element
[…]

### 2025-03-10 stoch_traj.m

@@ -309,6 +309,7 @@ plot_corr_2d(corr_2d, "Correlation between costs in $J_h(u_h)$ and $J_l(u_{hla})
 
 %% F num opt with CV estimator
 cv = Cv(x0_ext_mean, x0_ext_cov, data.lqrsol{1}, data.lqrsol{2});
+acv = Acv(x0_ext_mean, x0_ext_cov, data.lqrsol{1}, data.lqrsol{2});
 %% F.1 normal CV Estimator
 cv_samples = rv_samples/2;
 [~, U_hf, U_lf] = cv.opt(u0_num, -1, -1, x0_rv_ext, cv_samples, false);
@@ -556,13 +557,12 @@ for num_samples=num_rv_samples
     x0_rv_ext = [x0_rv; repmat(u0, 1, n_total); repmat(ref, 1, n_total)];
     
     % CV with fixed LF solution
-    % [costs, Us] = Est.CvOpt(u0_num, max_iters, x0_rv_ext, x0_ext_mean, x0_ext_cov, data.lqrsol{1}, data.lqrsol{2}, n_cv, true);
     [costs, Us, ~] = cv.opt(u0_num, max_iters, tol, x0_rv_ext, n_cv, true);
     data.cv_fix_cost(:, i, num_rv_samples == num_samples) = costs;
     data.cv_fix_u(:, :, i, num_rv_samples == num_samples) = Us;
     
     % CV
-    % [costs, Us] = Est.CvOpt(u0_num, max_iters, x0_rv_ext, x0_ext_mean, x0_ext_cov, data.lqrsol{1}, data.lqrsol{2}, n_cv, false);
+    % [costs, Us, ~] = cv.opt(u0_num, max_iters, tol, x0_rv_ext, n_cv, false);
     % data.cv_cost(:, i, num_rv_samples == num_samples) = costs;
     % data.cv_u(:, :, i, num_rv_samples == num_samples) = Us;
     
@@ -596,12 +596,12 @@ for num_samples=num_rv_samples
     x0_rv_ext = [x0_rv; repmat(u0, 1, n_total); repmat(ref, 1, n_total)];
     
     % ACV with fixed LF solution
-    [costs, Us] = Est.AcvOpt(u0_num, max_iters, x0_rv_ext, data.lqrsol{1}, data.lqrsol{2}, n_acv, m_acv, true);
[…]

### 2025-03-10 Mc.m
cleanup opt sim, add class for mc

@@ -1,0 +1,67 @@
+classdef Mc < handle
+	properties
+		Us = [];
+		U_hlas = [];
+		idx = 1;
+		x0_mean = [];
+		x0_cov = [];
+		lqrsol = {};
+	end
+	
+	methods
+		function obj = Mc(x0_mean, x0_cov, lqrsol)
+			obj.x0_mean = x0_mean;
+			obj.x0_cov = x0_cov;
+			obj.lqrsol = lqrsol;
+		end
+		
+		function cost = est(obj, x0_rv_ext, n, u)
+			cost = mean(St.LQRCost(x0_rv_ext(:, 1:n), obj.lqrsol, u));
+		end
+		
+		function [costs, Us, U_hlas] = opt(obj, u0, max_iters, tol, x0_rv_ext, n)
+			costs = zeros(max_iters, 1);
+			obj.Us = zeros(size(u0, 1), max_iters);
+			obj.U_hlas = zeros(size(u0, 1)/10, max_iters);
+			obj.idx = 1;
+			
+			if max_iters <  0 || tol < 0
+				options = optimoptions('fminunc', 'OutputFcn', @OutFn);
+			else
+				options = optimoptions('fminunc', 'OutputFcn', @OutFn, 'MaxIter', max_iters, 'OptimalityTolerance', tol);
+			end
+			fminunc(@(u) obj.est(x0_rv_ext, n, u), u0, options);
+			
+			% trim to match iters
+			obj.idx = obj.idx-1; % remove last iteration that stopped it
+			if max_iters < 0
+				iters = obj.idx;
+			else
+				iters = min(obj.idx, max_iters);
+			end
+			costs = costs(1:iters);
+			obj.Us = obj.Us(:, 1:iters);
+			obj.U_hlas = Est.DownsampleAvg(obj.Us, 10);
+			
+			function stop = OutFn(x, optimValues, state)
+				stop = false;
+				if isequal(state, 'iter')
+					% obj.idx starts at 1, so increment after assignment
+					obj.Us(:, obj.idx) = x;
+					costs(obj.idx) = optimValues.fval;
+					obj.idx = obj.idx+1;
+				end
+			end
+			Us = obj.Us;
+			U_hlas = obj.U_hlas;
+		end
+		
+		function var = var(n, u)
+			var_h = St.LQRVar(obj.x0_mean, obj.x0_cov, obj.lqrsol, u);
+			var = var_h/n;
+		end
+		
+		
+	end
+	
+end

### 2025-03-10 stoch_traj.m

@@ -204,7 +204,7 @@ for idx = 1:length(TsList)
   saveas(fig, "figs/"+Ts+"_cost_dist.svg");
 end
 
-% *** E. correlation check for CV, ONLY FOR 2 LEVEL TODO: extend to work for any number of levels
+% *** E. correlation check for CV
 % E.1 correlation with perturbation in direction of maximum ascent
 
 %% E.1.1 perturbation in HF objective fn
@@ -281,7 +281,6 @@ cost_hf = St.LQRCostMulti(x0_rv_ext, data.lqrsol{1}, U_hf);
 cost_lf = St.LQRCostMulti(x0_rv_ext, data.lqrsol{2}, U_lf);
 corr_st = St.CorrMulti(cost_hf, cost_lf);
 corr_an = St.LQRCorrMulti(x0_ext_mean, x0_ext_cov, data.lqrsol{1}, data.lqrsol{2}, U_hf, U_lf);
-% corr_2d = St.CorrMulti2D(cost_hf, cost_lf);
 corr_2d = St.LQRCorrMulti2D(x0_ext_mean, x0_ext_cov, data.lqrsol{1}, data.lqrsol{2}, U_hf, U_lf);
 
 % plot correlation
@@ -298,7 +297,6 @@ cost_hf = St.LQRCostMulti(x0_rv_ext, data.lqrsol{1}, U_hf);
 cost_lf = St.LQRCostMulti(x0_rv_ext, data.lqrsol{2}, U_lf);
 corr_st = St.CorrMulti(cost_hf, cost_lf);
 corr_an = St.LQRCorrMulti(x0_ext_mean, x0_ext_cov, data.lqrsol{1}, data.lqrsol{2}, U_hf, U_lf);
-% corr_2d = St.CorrMulti2D(cost_hf, cost_lf);
 corr_2d = St.LQRCorrMulti2D(x0_ext_mean, x0_ext_cov, data.lqrsol{1}, data.lqrsol{2}, U_hf, U_lf);
 
 % plot correlation
@@ -309,7 +307,6 @@ plot_corr_2d(corr_2d, "Correlation between costs in $J_h(u_h)$ and $J_l(u_{hla})
 
 %% F num opt with CV estimator
 cv = Cv(x0_ext_mean, x0_ext_cov, data.lqrsol{1}, data.lqrsol{2});
[…]

### 2025-03-10 artefacts/paper.tex
write motivation and add experiments to extra section

@@
 
-The objective of this paper is to reduce the variance of the objective function estimation under sample average approximation.
-A Monte Carlo(MC) simulation can be used to estimate the expected value of the objective function, but a control variate(CV) approach
-would reduce the variance with the same computational cost. In the absence of an analytical expression of the expected value of the
-objective function, an approximate control variate(ACV) approach is also proposed. The CV estimator will also serve as
-the benchmark for the ACV estimator, since it is the best possible variance reduction that can be achieved with the ACV estimator.
+For optimizing the trajectory of system under uncertainty, a common approach is to compute the expectation of the objective
+function of a trajectory is by doing sample average approximation, where we propagate the trajectory with multiples samples
+of the random variable, and take the average of all samples in a numerical optimizer to determine the optimal control inputs.
+The most straighforward way to compute the expectation of the trajectory's objective function is to use Monte Carlo (MC) estimator.
+When using a MC estimator, the variance of the estimated trajectory objective is related to the variance of the trajectory's
+objective function in the following way.
 
+$$
+\text{Var}[S_n^{MC}(J)] = \frac{\text{Var}[J]}{n}
+$$
+
+Where $S_n^{MC}(J)$ is the MC estimator of the objective function $J$ with $n$ samples of the random variable.
 The solution to the stochastic LQR controller, $U^*$ can be written as a function of the initial state $x_0$.
@@
 
-Thus, we aim to reduce the variance of the objective function estimate thereby reducing the covariance of the
-solution $U^*$ given a finite cost of computation.
+The objective of this paper is to reduce the variance of the objective function estimation under sample average approximation,
+thus, thereby reducing the covariance of the solution $U^*$ given a finite cost of computation.
+
+A control variate (CV) estimator is a variance reduction technique that uses the correlation between the objective function
+and another function to reduce the variance of the objective function estimation. Its variance has the form
+
+$$
+\text{Var}[S_{n_{cv}}^{CV}(J_h, J_l)] = \text{Var}[J_h] (1 - \rho^2)
+$$
+
+Where $J_h$ and $J_l$ are the high and low fidelity models of the objective function, $\rho$ is the Pearson correlation
+coefficient between the high and low fidelity models, and $S_{n_{cv}}^{CV}(J_h, J_l)$ is the CV estimator of the objective
+function $J_h$ using the low fidelity model $J_l$.
+
+The CV estimator requires the analytical expectation of the expectation of the low fidelity model, and the covariance between
+the high and low fidelity models. This might not be available in practice, and hence, an alternative is the approximate
+control variate (ACV) estimator. The ACV estimator uses the statistical estimate of the expectation and covariance
+of the low fidelity model, and has the variance
+
+$$
+\text{Var}[S_{n_{acv},m_{acv}}^{ACV}(J_h, J_l)] = \text{Var}[J_h] (1 - \frac{m_{acv}}{m_{acv} + n_{acv}} \rho^2)
+$$
+
+Where $m_{acv}$ and $n_{acv}$ are the number of samples used to estimate the expectation and covariance of the low fidelity model,
 
@@
 
+\section{Extra}
+To test the estimators, we set up a couple of experiments.
+
+We fixed $\alpha = -1$ for the ACV estimator and increased the ratio $m:n$ for the ACV estimator
+to see what the optimal sample allocation is
+\begin{algorithm}
+  \caption{ACV sample allocation experiment}
+  \begin{algorithmic}
+    \Require $J_h, J_l, x_0, e_s, ratio_{cost}, MC_{cost}, ratio_{mn}$
+    \State $x_{0,samples} \gets \text{generate samples of } x_0 \text{ for all } e_s$
+    \For{$r$ in $ratio_{mn}$}
+    \State $n_{acv} \gets MC_{cost} / (1 + ratio_{cost} + r * ratio_{cost})$
+    \State $m_{acv} \gets r * n_{acv}$
+    \For{$x_0^{(i)}$ in $x_{0,samples}$}
+    \State $O^{(i)} \gets S_{n_{acv},m_{acv}}^{ACV}(J_h, J_l, x_0^{(i)})$
+    \EndFor
+    \State $\text{Var}^{(r)} \gets \text{Var}[O]$
+    \EndFor
+  \end{algorithmic}
+\end{algorithm}
+
+$e_s$ is the number of estimator samples
+
+$ratio_{cost}$ is the ratio of the cost of $J_l$ to $J_h$
+
+$MC_{cost}$ is the equivalent cost of the MC estimator
+
+Then we fixed $n$ for the ACV estimator and increased $m$ to see if it converges to the CV estimator variance
+\begin{algorithm}
+  \caption{ACV convergence experiment}
+  \begin{algorithmic}
+    \Require $J_h, J_l, x_0, e_s, ratio_{cost}, MC_{cost}, ms_{acv}$
+    \State $x_{0,samples} \gets \text{generate samples of } x_0 \text{ for all } e_s$
+    \State $n_{cv} \gets MC_{cost} / (1 + ratio_{cost})$
+    \State $n_{acv} \gets MC_{cost} / (1 + ratio_{cost})$ (same as CV)
+    \For{$m$ in $ms_{acv}$}
+    \State $m_{acv} \gets m$
+    \For{$x_0^{(i)}$ in $x_{0,samples}$}
+    \State $O_{cv}^{(i)} \gets S_{n_{cv}}^{CV}(J_h, J_l, x_0^{(i)})$
+    \State $O_{acv}^{(i)} \gets S_{n_{acv},m_{acv}}^{ACV}(J_h, J_l, x_0^{(i)})$
+    \EndFor
+    \State $\text{Var}_{cv}^{(m)} \gets \text{Var}[O_{cv}]$
+    \State $\text{Var}_{acv}^{(m)} \gets \text{Var}[O_{acv}]$
+    \State $\text{Var}^{(m)} \gets (\text{Var}_{acv}^{(m)} - \text{Var}_{cv}^{(m)})^2$
+    \EndFor
+  \end{algorithmic}
+\end{algorithm}
+
 \end{document}
